---
title: "survival_0.3_fixed"
output: html_notebook
---


```{r}
library(tidyverse)
library(survival)
library(survminer)
library(dplyr)
library(ggplot2)
library(stringr)

```


#Read in files 
```{r}

# Load input files
clin_pt  <- read.delim("data_clinical_patient.txt", skip = 4, check.names = FALSE)
clin_smp <- read.delim("data_clinical_sample.txt", skip = 4, check.names = FALSE)

# Load CNV table and clean SAMPLE_ID
cnv_raw <- read.delim("all_cnv_table.txt", check.names = FALSE, quote = "\"") %>%
  tibble::rownames_to_column("SAMPLE_ID")

clin_smp$SAMPLE_ID <- as.character(clin_smp$SAMPLE_ID)
cnv_raw$SAMPLE_ID  <- as.character(cnv_raw$SAMPLE_ID)

cytoband_cols <- grep("^[0-9XY]+[pq]", names(cnv_raw), value = TRUE)
arm_cols <- grep("^Arm", names(cnv_raw), value = TRUE)

cnv_cat <- cnv_raw %>%
  mutate(across(all_of(c(cytoband_cols, arm_cols)), ~ case_when(
    .x >  0.3 ~ "Gain",
    .x < -0.3 ~ "Loss",
    TRUE      ~ "None"
  )))


## MERGE
# Step 1: Prepare clean sample info
sample_info <- clin_smp %>%
  select(SAMPLE_ID, PATIENT_ID, CANCER_TYPE, ONCOTREE_CODE, SAMPLE_TYPE)

# Step 2: Join sample_info with CNV matrix
merged_sample <- sample_info %>%
  left_join(cnv_cat, by = "SAMPLE_ID")

# Step 3: Join patient-level clinical data
df <- clin_pt %>%
  select(PATIENT_ID, OS_MONTHS, OS_STATUS, SEX, DRUG_TYPE) %>%
  left_join(merged_sample, by = "PATIENT_ID") %>%
  mutate(OS = if_else(str_detect(OS_STATUS, "LIVING"), 0, 1))

length(unique(df$PATIENT_ID))
```


# KM curves
## cancer types 
```{r}
###### all cancers 
# Initialize table of filtered hits
output_dir <- "KM_plots_PD1_0.3_cancertype"
km_sig_hits <- data.frame()

# Loop through all cancer types directly
for (tumor in unique(df$CANCER_TYPE)) {
  tumor_data <- df %>%
    filter(CANCER_TYPE == tumor, DRUG_TYPE == "PD-1/PDL-1")
  
  for (arm in grep("^Arm", names(tumor_data), value = TRUE)) {
    for (comparison in c("Gain", "Loss")) {
      # Get the specific arm column values
      arm_values <- tumor_data[[arm]]
      
      # Filter for only patients with "None" or the specific comparison
      subset <- tumor_data %>%
        filter(!!sym(arm) == "None" | !!sym(arm) == comparison) %>%
        mutate(CNVstatus = factor(!!sym(arm), levels = c(comparison, "None")))
      
      if (nrow(subset) >= 10 && length(unique(subset$CNVstatus)) == 2) {
        # Log-rank test
        surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
        surv_diff <- survdiff(surv_obj ~ CNVstatus, data = subset)
        km_p <- 1 - pchisq(surv_diff$chisq, df = 1)
        
        if (km_p < 0.1) {
          # Save to table and plot
          km_sig_hits <- rbind(km_sig_hits, data.frame(
            CANCER_TYPE = tumor,
            Arm = arm,
            Comparison = comparison,
            KMP = signif(km_p, 3),
            N = nrow(subset)
          ))
          
          # Create plot
          fit <- survfit(Surv(OS_MONTHS, OS) ~ CNVstatus, data = subset)
          
          # Choose color palette depending on comparison type
          if (comparison == "Gain") {
            current_palette <- c("#EA6A6A", "black")  # Gain = red, None = black
          } else if (comparison == "Loss") {
            current_palette <- c("#4F6DB8", "black")  # Loss = blue, None = black
          }
          
          # Plot
          p <- ggsurvplot(fit, data = subset,
                          pval = TRUE,
                          risk.table = TRUE,
                          conf.int = FALSE,
                          palette = current_palette,
                          title = paste(tumor, arm, comparison, "vs None – PD-1/PDL-1"),
                          xlab = "Survival Time (months)",
                          ylab = "Survival probability",
                          risk.table.height = 0.25,
                          break.time.by = 20,
                          ggtheme = theme_classic(base_size = 14),
                          linetype = "solid",
                          size = 1.4)
    
          # Save plot
          filename <- paste0(output_dir, "/", gsub(" ", "_", tumor), "_", arm, "_", comparison, "_PD1.pdf")
          # Create combined plot with risk table
          combined_plot <- arrangeGrob(p$plot, p$table, ncol = 1, heights = c(3, 1))
          ggsave(filename, plot = combined_plot, width = 7, height = 7)
        }
      }
    }
  }
}

```

## Oncotree
```{r}
###### all cancers 
# Initialize table of filtered hits
output_dir <- "KM_plots_PD1_0.3_oncotree"
km_sig_hits <- data.frame()

# Loop through all cancer types directly
for (tumor in unique(df$ONCOTREE_CODE)) {
  tumor_data <- df %>%
    filter(ONCOTREE_CODE == tumor, DRUG_TYPE == "PD-1/PDL-1")
  
  for (arm in grep("^Arm", names(tumor_data), value = TRUE)) {
    for (comparison in c("Gain", "Loss")) {
      # Get the specific arm column values
      arm_values <- tumor_data[[arm]]
      
      # Filter for only patients with "None" or the specific comparison
      subset <- tumor_data %>%
        filter(!!sym(arm) == "None" | !!sym(arm) == comparison) %>%
        mutate(CNVstatus = factor(!!sym(arm), levels = c(comparison, "None")))
      
      if (nrow(subset) >= 10 && length(unique(subset$CNVstatus)) == 2) {
        # Log-rank test
        surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
        surv_diff <- survdiff(surv_obj ~ CNVstatus, data = subset)
        km_p <- 1 - pchisq(surv_diff$chisq, df = 1)
        
        if (km_p < 0.1) {
          # Save to table and plot
          km_sig_hits <- rbind(km_sig_hits, data.frame(
            CANCER_TYPE = tumor,
            Arm = arm,
            Comparison = comparison,
            KMP = signif(km_p, 3),
            N = nrow(subset)
          ))
          
          # Create plot
          fit <- survfit(Surv(OS_MONTHS, OS) ~ CNVstatus, data = subset)
          
          # Choose color palette depending on comparison type
          if (comparison == "Gain") {
            current_palette <- c("#EA6A6A", "black")  # Gain = red, None = black
          } else if (comparison == "Loss") {
            current_palette <- c("#4F6DB8", "black")  # Loss = blue, None = black
          }
          
          # Plot
          p <- ggsurvplot(fit, data = subset,
                          pval = TRUE,
                          risk.table = TRUE,
                          conf.int = FALSE,
                          palette = current_palette,
                          title = paste(tumor, arm, comparison, "vs None – PD-1/PDL-1"),
                          xlab = "Survival Time (months)",
                          ylab = "Survival probability",
                          risk.table.height = 0.25,
                          break.time.by = 20,
                          ggtheme = theme_classic(base_size = 14),
                          linetype = "solid",
                          size = 1.4)
    
          # Save plot
          filename <- paste0(output_dir, "/", gsub(" ", "_", tumor), "_", arm, "_", comparison, "_PD1.pdf")
          # Create combined plot with risk table
          combined_plot <- arrangeGrob(p$plot, p$table, ncol = 1, heights = c(3, 1))
          ggsave(filename, plot = combined_plot, width = 7, height = 7)
        }
      }
    }
  }
}


```


## cytobands associated with poor prongnonosis 
```{r}
# grep cytoabnd level data 
cytoband_cols <- grep("^[0-9XY]+[pq]", names(df), value = TRUE)

km_sig_hits <- data.frame()

for (tumor in unique(df$CANCER_TYPE)) {
  tumor_data <- df %>% filter(CANCER_TYPE == tumor, DRUG_TYPE == "PD-1/PDL-1")
  
  for (band in cytoband_cols) {
    for (comparison in c("Gain", "Loss")) {
      subset <- tumor_data %>%
        filter(!!sym(band) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(band), levels = c(comparison, "None")))
      
      if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
        surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
        surv_diff <- survdiff(surv_obj ~ CNV, data = subset)
        km_p <- 1 - pchisq(surv_diff$chisq, df = 1)
        
        if (km_p < 0.1) {
          km_sig_hits <- rbind(km_sig_hits, data.frame(
            CANCER_TYPE = tumor,
            Cytoband = band,
            Comparison = comparison,
            KMP = signif(km_p, 3),
            N = nrow(subset)
          ))
        }
      }
    }
  }
}

write.csv(km_sig_hits, "significant_survival_cytobands.csv", row.names = FALSE)
```

oncotree
```{r}
library(tidyverse)
library(survival)

# Get cytoband-level CNV columns
cytoband_cols <- grep("^[0-9XY]+[pq]", names(df), value = TRUE)

# Output container
km_sig_hits <- data.frame()

# Loop over OncoTree codes (not CANCER_TYPE)
for (tumor in unique(df$ONCOTREE_CODE)) {
  tumor_data <- df %>% filter(ONCOTREE_CODE == tumor, DRUG_TYPE == "PD-1/PDL-1")

  for (band in cytoband_cols) {
    for (comparison in c("Gain", "Loss")) {
      subset <- tumor_data %>%
        filter(!!sym(band) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(band), levels = c(comparison, "None")))

      if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
        surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
        surv_diff <- survdiff(surv_obj ~ CNV, data = subset)
        km_p <- 1 - pchisq(surv_diff$chisq, df = 1)

        if (km_p < 0.1) {
          km_sig_hits <- rbind(km_sig_hits, data.frame(
            OncoTree = tumor,
            Cytoband = band,
            Comparison = comparison,
            KMP = signif(km_p, 3),
            N = nrow(subset)
          ))
        }
      }
    }
  }
}

# Save result to CSV
write.csv(km_sig_hits, "significant_survival_cytobands_OncoTree.csv", row.names = FALSE)
```

### melanoma
```{r}
 # Filter for SKCM + significant cytobands on 1q or 9p
skcm_cytoband_hits <- km_sig_hits %>%
  filter(CANCER_TYPE == "Melanoma",
         grepl("^1q|^9p", Cytoband),
         KMP < 0.1) %>%
  arrange(Cytoband)

# View results
print(skcm_cytoband_hits, n = Inf)

# Optional: save to file
write.csv(skcm_cytoband_hits, "SKCM_significant_1q_9p_cytobands.csv", row.names = FALSE)
```









# HR curves 
```{r}
###### HR curves
library(tidyverse)
library(gtools)  # for mixedsort
library(survival)

# First get cytoband columns (again) to be safe
cytoband_cols <- grep("^[0-9XY]+[pq]", names(df), value = TRUE)

results_cytoband <- data.frame()

for (tumor in unique(df$CANCER_TYPE)) {
  tumor_data <- df %>% filter(CANCER_TYPE == tumor, DRUG_TYPE == "PD-1/PDL-1")
  
  for (band in cytoband_cols) {
    for (comparison in c("Gain", "Loss")) {
      subset <- tumor_data %>%
        filter(!!sym(band) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(band), levels = c("None", comparison)))
      
      if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
        tryCatch({
          fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV, data = subset)
          hr <- exp(coef(fit))
          ci <- round(exp(confint(fit)), 3)
          p <- summary(fit)$coefficients[5]
          
          results_cytoband <- rbind(results_cytoband, data.frame(
            CANCER_TYPE = tumor,
            Cytoband = band,
            Comparison = comparison,
            HR = round(hr, 3),
            CI = paste(ci[1], ci[2], sep = " - "),
            P = signif(p, 3),
            log2HR = log2(hr),
            Event = comparison,
            N = nrow(subset)
          ))
        }, error = function(e){})
      }
    }
  }
}


# Reorder Cytoband column factor levels in natural chromosomal order
results_cytoband$Cytoband <- factor(
  results_cytoband$Cytoband,
  levels = mixedsort(unique(results_cytoband$Cytoband))
)

# Rename P to something more descriptive
results_cytoband <- results_cytoband %>%
  rename(Raw_P_Cox = P)

# Arrange by tumor type, then cytoband, then gain/loss
results_cytoband <- results_cytoband %>%
  arrange(CANCER_TYPE, Cytoband, Event)

# Save updated version
write.csv(results_cytoband, "All_CancerTypes_Cytoband_SurvivalResults.csv", row.names = FALSE)
```


```{r}
### Graph  
# Step 4: Axis label mapping — 1 label per arm
axis_labels <- results_cytoband %>%
  distinct(Cytoband, Arm) %>%
  group_by(Arm) %>%
  arrange(Cytoband, .by_group = TRUE) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(ArmLabel = Arm)

# Step 5: Cytoband ordering
results_cytoband$Cytoband <- factor(
  results_cytoband$Cytoband,
  levels = mixedsort(unique(results_cytoband$Cytoband))
)

# Step 6: Ensure all combinations show up (even if missing)
for (tumor in unique(results_cytoband$CANCER_TYPE)) {
  combo <- expand.grid(
    Cytoband = levels(results_cytoband$Cytoband),
    Event = c("Gain", "Loss")
  )
  combo$CANCER_TYPE <- tumor
  
  plot_data <- combo %>%
    left_join(results_cytoband %>% filter(CANCER_TYPE == tumor),
              by = c("Cytoband", "Event", "CANCER_TYPE")) %>%
    mutate(
      sign_logP = replace_na(sign_logP, 0),
      Arm = str_extract(Cytoband, "^[0-9XY]+[pq]")
    )
  
  p <- ggplot(plot_data, aes(x = Cytoband, y = sign_logP, fill = Event)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.8) +
    geom_hline(yintercept = c(-1.3, 1.3), linetype = "dotted", color = "grey40") +
    geom_hline(yintercept = 0, linetype = "solid") +
    scale_fill_manual(values = c("Gain" = "#D64045", "Loss" = "#375E97")) +
    scale_x_discrete(
      breaks = axis_labels$Cytoband,
      labels = axis_labels$ArmLabel
    ) +
    labs(
      title = paste0("Signed -log10(p) CNV Survival – ", tumor),
      x = "Chromosome Arm",
      y = "Signed -log10(p)",
      fill = "CNV Event"
    ) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
      plot.title = element_text(face = "bold", size = 12)
    )
  
  ggsave(
    filename = paste0("HR_logP_0.3/", gsub(" ", "_", tumor), "_CNV_cytoband_plot.pdf"),
    plot = p, width = 12, height = 4
  )
}

```

## Adjust schematics
```{r}
# Filter for Melanoma (in CANCER_TYPE)
tumor <- "Melanoma"
plot_data <- results_cytoband %>%
  filter(CANCER_TYPE == tumor) %>%
  mutate(
    Arm = str_extract(Cytoband, "^[0-9XY]+[pq]"),
    Chromosome = str_extract(Arm, "^[0-9XY]+")
  )

# Ensure Cytoband order is preserved
plot_data$Cytoband <- factor(plot_data$Cytoband, levels = mixedsort(unique(plot_data$Cytoband)))

# Get centromere boundaries
centromere_breaks <- plot_data %>%
  filter(str_detect(Arm, "q")) %>%
  group_by(Chromosome) %>%
  slice(1) %>%
  ungroup()

# Plot
p <- ggplot(plot_data, aes(x = Cytoband, y = sign_logP, fill = Event)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.8) +
  geom_hline(yintercept = c(-1.3, 1.3), linetype = "dotted", color = "grey40") +
  geom_hline(yintercept = 0, linetype = "solid") +
  scale_fill_manual(values = c("Gain" = "#D64045", "Loss" = "#375E97")) +
  scale_x_discrete(
    breaks = centromere_breaks$Cytoband,
    labels = centromere_breaks$Chromosome
  ) +
  labs(
    title = "Signed -log10(p) CNV Survival – Melanoma",
    x = "chromosome",
    y = "signed -log(p)",
    fill = "Event"
  ) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),  # remove grid
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 6),
    plot.title = element_text(face = "bold", size = 12)
  )

# Add thin dashed vertical centromere lines
for (cb in centromere_breaks$Cytoband) {
  p <- p + geom_vline(xintercept = which(levels(plot_data$Cytoband) == cb),
                      linetype = "dashed", color = "black", alpha = 0.3, linewidth = 0.3)
}

print(p)

```

cancer type 
```{r}
library(tidyverse)
library(gtools)

# Loop over each tumor type
for (tumor in unique(results_cytoband$CANCER_TYPE)) {
  plot_data <- results_cytoband %>%
    filter(CANCER_TYPE == tumor) %>%
    mutate(
      Arm = str_extract(Cytoband, "^[0-9XY]+[pq]"),
      Chromosome = str_extract(Arm, "^[0-9XY]+")
    )
  
  plot_data$Cytoband <- factor(plot_data$Cytoband, levels = mixedsort(unique(plot_data$Cytoband)))
  
  # Get centromere breakpoints
  centromere_breaks <- plot_data %>%
    filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>%
    slice(1) %>%
    ungroup()
  
  p <- ggplot(plot_data, aes(x = Cytoband, y = sign_logP, fill = Event)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.8) +
    geom_hline(yintercept = c(-1.3, 1.3), linetype = "dotted", color = "grey40") +
    geom_hline(yintercept = 0, linetype = "solid") +
    scale_fill_manual(values = c("Gain" = "#D64045", "Loss" = "#375E97")) +
    scale_x_discrete(
      breaks = centromere_breaks$Cytoband,
      labels = centromere_breaks$Chromosome
    ) +
    labs(
      title = paste("Signed -log10(p) CNV Survival –", tumor),
      x = "chromosome",
      y = "signed -log(p)",
      fill = "Event"
    ) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(size = 6),
      plot.title = element_text(face = "bold", size = 12)
    )
  
  for (cb in centromere_breaks$Cytoband) {
    p <- p + geom_vline(xintercept = which(levels(plot_data$Cytoband) == cb),
                        linetype = "dashed", color = "black", alpha = 0.3, linewidth = 0.3)
  }
  
  # Save to file
  ggsave(
    filename = paste0("HR_logP_0.3/", gsub(" ", "_", tumor), "_CNV_cytoband_plot.pdf"),
    plot = p, width = 12, height = 4
  )
}
```


Oncotree
```{r}
# Apply schematics to ONCOTREE_CODE instead of CANCER_TYPE
results_cytoband <- rbind(results_cytoband, data.frame(
  CANCER_TYPE = tumor,
  ONCOTREE_CODE = unique(tumor_data$ONCOTREE_CODE)[1],  # grab the unique ONCOTREE code
  Cytoband = band,
  Comparison = comparison,
  HR = round(hr, 3),
  CI = paste(ci[1], ci[2], sep = " - "),
  P = signif(p, 3),
  log2HR = log2(hr),
  Event = comparison,
  N = nrow(subset)
))

# Loop over each OncoTree code
for (tumor in unique(results_cytoband$ONCOTREE_CODE)) {
  plot_data <- results_cytoband %>%
    filter(ONCOTREE_CODE == tumor) %>%
    mutate(
      Arm = str_extract(Cytoband, "^[0-9XY]+[pq]"),
      Chromosome = str_extract(Arm, "^[0-9XY]+")
    )
  
  if (nrow(plot_data) == 0) next  # skip if empty
  
  plot_data$Cytoband <- factor(plot_data$Cytoband, levels = mixedsort(unique(plot_data$Cytoband)))
  
  centromere_breaks <- plot_data %>%
    filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>%
    slice(1) %>%
    ungroup()
  
  p <- ggplot(plot_data, aes(x = Cytoband, y = sign_logP, fill = Event)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.8) +
    geom_hline(yintercept = c(-1.3, 1.3), linetype = "dotted", color = "grey40") +
    geom_hline(yintercept = 0, linetype = "solid") +
    scale_fill_manual(values = c("Gain" = "#D64045", "Loss" = "#375E97")) +
    scale_x_discrete(
      breaks = centromere_breaks$Cytoband,
      labels = centromere_breaks$Chromosome
    ) +
    labs(
      title = paste("Signed -log10(p) CNV Survival –", tumor),
      x = "chromosome",
      y = "signed -log(p)",
      fill = "Event"
    ) +
    theme_bw() +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(size = 6),
      plot.title = element_text(face = "bold", size = 12)
    )
  
  for (cb in centromere_breaks$Cytoband) {
    p <- p + geom_vline(xintercept = which(levels(plot_data$Cytoband) == cb),
                        linetype = "dashed", color = "black", alpha = 0.3, linewidth = 0.3)
  }
  
  ggsave(
    filename = paste0("HR_logP_0.3_oncotree/", gsub(" ", "_", tumor), "_CNV_cytoband_plot.pdf"),
    plot = p, width = 12, height = 4
  )
}
```


## area plots
```{r}
library(dplyr)
library(ggplot2)
library(gtools)   # mixedsort()

tumor <- "Melanoma"
chr_order <- c(as.character(1:22), "X", "Y")   # desired order

plot_data <- results_cytoband %>%
  filter(CANCER_TYPE == tumor) %>%
  mutate(
    Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
    Chromosome = str_extract(Arm, "^[0-9XY]+"),
    Chromosome = factor(Chromosome, levels = chr_order)        # label by order 
  ) %>%
  arrange(Chromosome, Cytoband) %>%                            # now numeric order
  mutate(
    Cytoband = factor(Cytoband, levels = mixedsort(unique(Cytoband))),
    idx      = row_number()
  )

## guides
centromere_idx <- plot_data %>%            # first q-band in each chromosome
  filter(str_detect(Arm, "q")) %>% 
  group_by(Chromosome) %>% 
  slice_head(n = 1) %>% 
  pull(idx)

chr_last_idx <- plot_data %>%              # last band in each chromosome
  group_by(Chromosome) %>% 
  slice_tail(n = 1) %>% 
  pull(idx)

chr_mid_idx <- plot_data %>%               # mid-point for x-axis label
  group_by(Chromosome) %>% 
  summarise(idx = floor(mean(range(idx))), .groups = "drop")

## plot
p <- ggplot(plot_data, aes(idx, sign_logP, fill = Event, group = Event)) +
  geom_area(position = "identity", alpha = 0.75) +
  geom_hline(yintercept = 0,            linewidth = .25, colour = "black") +
  geom_hline(yintercept = c(-1.3, 1.3), linewidth = .25,
             colour = "grey40", linetype = "dotted") +
  geom_vline(xintercept = chr_last_idx + 0.5, colour = "black", linewidth = .3) +
  geom_vline(xintercept = centromere_idx,     colour = "grey30",
             linewidth = .25, linetype = "dashed") +
  scale_fill_manual(values = c(Gain = "#D64045", Loss = "#375E97")) +
  scale_x_continuous(
    breaks = chr_mid_idx$idx,
    labels = chr_mid_idx$Chromosome,
    expand = c(0, 0)
  ) +
  labs(
    title = "Signed -log10(p) CNV Survival – Melanoma",
    x     = "chromosome",
    y     = "signed -log(p)",
    fill  = "Event"
  ) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x      = element_text(size = 6),
    plot.title       = element_text(face = "bold", size = 12)
  )

print(p)
```

```{r}
library(dplyr)
library(ggplot2)
library(gtools)
library(stringr)

tumor <- "Bladder Cancer"                            
chr_order <- c(as.character(1:22), "X", "Y")   # karyotypic order

plot_data <- results_cytoband %>% 
  filter(CANCER_TYPE == tumor) %>% 
  mutate(
    Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
    Chromosome = factor(str_extract(Arm, "^[0-9XY]+"), levels = chr_order)
  ) %>% 
  arrange(Chromosome, Cytoband) %>% 
  mutate(
    Cytoband = factor(Cytoband, levels = mixedsort(unique(Cytoband))),
    idx      = row_number()
  )

# positions for guides
centromere_idx <- plot_data %>% 
  filter(str_detect(Arm, "q")) %>% 
  group_by(Chromosome) %>% 
  slice_head(n = 1) %>% pull(idx)

chr_last_idx <- plot_data %>% 
  group_by(Chromosome) %>% 
  slice_tail(n = 1) %>% pull(idx)

chr_mid_idx <- plot_data %>% 
  group_by(Chromosome) %>% 
  summarise(idx = floor(mean(range(idx))), .groups = "drop")

# plot
p <- ggplot(plot_data, aes(idx, sign_logP, fill = Event)) +
  geom_area(position = "identity", alpha = 0.75) +                             # silhouette
  geom_hline(yintercept = 0, linewidth = .25) +                                # baseline
  geom_hline(yintercept = c(-1.3, 1.3), colour = "grey40",                     # ±sig guide
             linetype = "dotted", linewidth = .25) +
  geom_vline(xintercept = chr_last_idx + 0.5,                                  # solid chr borders
             colour = "black", linewidth = .25) +
  geom_vline(xintercept = centromere_idx,                                      # dashed centromeres
             colour = "grey30", linetype = "dashed", linewidth = .15) +
  scale_fill_manual(values = c(Gain = "#D64045", Loss = "#375E97")) +
  scale_x_continuous(breaks = chr_mid_idx$idx,
                     labels = chr_mid_idx$Chromosome,
                     expand = c(0, 0)) +
  labs(title = "Signed -log10(p) CNV Survival – Bladder Cancer",
       x = "chromosome",
       y = "signed -log(p)",
       fill = "Event") +
  theme_bw() +
  theme(
    panel.border     = element_blank(),                # remove box
    panel.grid       = element_blank(),                # no background grid
    axis.line        = element_line(linewidth = .25),  # solid axes
    axis.text.x      = element_text(size = 9),
    axis.text.y      = element_text(size = 9),
    axis.title.x     = element_text(size = 11),
    axis.title.y     = element_text(size = 11, angle = 90),
    plot.title       = element_text(face = "bold", size = 12)
  )

print(p)
```

```{r}
## ─────────────────────────────────────────────────────────────────────────────
##  Signed –log10(p) CNV-survival “area” plots for every tumour type
##  -->  each plot is saved as HR_area_plots/<TUMOUR>_HR_area.pdf
## ─────────────────────────────────────────────────────────────────────────────

library(dplyr)
library(ggplot2)
library(stringr)
library(gtools)   # mixedsort()

# -----------------------------------------------------------------------------#
# 1.  Make sure an output directory exists
# -----------------------------------------------------------------------------#
out_dir <- "HR_area_plots"
if (!dir.exists(out_dir)) dir.create(out_dir)

# -----------------------------------------------------------------------------#
# 2.  Helper: build & return the ggplot object for one tumour
# -----------------------------------------------------------------------------#
plot_cnv_area <- function(df, tumour, signif_cut = 1.3) {
  
  chr_levels <- c(as.character(1:22), "X", "Y")
  
  pd <- df %>%                                             # karyotypic order
    filter(CANCER_TYPE == tumour) %>% 
    mutate(Arm        = str_extract(Cytoband, "^[0-9XY]+[pq]"),
           Chromosome = str_extract(Arm, "^[0-9XY]+"),
           Chromosome = factor(Chromosome, levels = chr_levels)) %>% 
    arrange(Chromosome, Cytoband) %>% 
    mutate(Cytoband = factor(Cytoband,
                             levels = mixedsort(unique(Cytoband))),
           idx      = row_number())
  
  # guides
  centromere_idx <- pd %>% filter(str_detect(Arm, "q")) %>%
    group_by(Chromosome) %>% slice_head(n = 1) %>% pull(idx)
  
  chr_last_idx   <- pd %>% group_by(Chromosome) %>%
    slice_tail(n = 1) %>% pull(idx)
  
  chr_mid_idx    <- pd %>% group_by(Chromosome) %>%
    summarise(idx = floor(mean(range(idx))), .groups = "drop")
  
  ggplot(pd, aes(idx, sign_logP, fill = Event)) +
    geom_area(position = "identity", alpha = 0.75) +
    geom_hline(yintercept = 0,            linewidth = .25) +
    geom_hline(yintercept = c(-signif_cut, signif_cut),
               colour = "grey40", linetype = "dotted", linewidth = .25) +
    geom_vline(xintercept = chr_last_idx + 0.5, colour = "black", linewidth = .3) +
    geom_vline(xintercept = centromere_idx, colour = "grey30",
               linetype = "dashed", linewidth = .25) +
    scale_fill_manual(values = c(Gain = "#D64045", Loss = "#375E97")) +
    scale_x_continuous(breaks = chr_mid_idx$idx,
                       labels = chr_mid_idx$Chromosome,
                       expand = c(0, 0)) +
    labs(title = paste("Signed -log10(p) CNV Survival –", tumour),
         x = "chromosome",
         y = "signed -log(p)",
         fill = "Event") +
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x      = element_text(size = 6),
          plot.title       = element_text(face = "bold", size = 12))
}

# -----------------------------------------------------------------------------#
# 3.  Loop over tumours, save one PDF each
# -----------------------------------------------------------------------------#
unique(results_cytoband$CANCER_TYPE) %>%
  sort() %>%                              # nicer alphabetical order
  purrr::walk(function(tumour) {
    p <- plot_cnv_area(results_cytoband, tumour)
    
    # tidy file-name: replace spaces / slashes with underscores
    fname <- gsub("[ /]", "_", tumour)
    ggsave(file.path(out_dir, paste0(fname, "_HR_area.pdf")),
           p, width = 10, height = 5, useDingbats = FALSE)
  })

```

# Look into subtypes 
First check where all the informations are stored
- Primary vs Metastasis: in SAMPLE_TYPE
- tumor subtypes in CANCER_TYPE_DETAILED 
```{r}
# Check column names and a few rows
names(clin_smp)
head(clin_smp, 10)
```

```{r}
table(df$SAMPLE_TYPE)
```

```{r}
# Filter for lung cancer and glioma cancer types
lung_glioma <- clin_smp %>%
  filter(CANCER_TYPE %in% c("Non-Small Cell Lung Cancer", "Glioma")) %>%
  group_by(CANCER_TYPE, CANCER_TYPE_DETAILED, ONCOTREE_CODE) %>%
  summarise(n_patients = n_distinct(PATIENT_ID), .groups = "drop")

print(lung_glioma)
```


### primary vs metastasis 
```{r}
library(tidyverse)
library(survival)

arm_cols <- grep("^Arm", names(df), value = TRUE)
primary_met_results <- data.frame()

for (tumor in unique(df$CANCER_TYPE)) {
  tumor_df <- df %>% filter(CANCER_TYPE == tumor)

  for (arm in arm_cols) {
    for (comparison in c("Gain", "Loss")) {
      
      # PRIMARY
      df_primary <- tumor_df %>%
        filter(SAMPLE_TYPE == "Primary", !!sym(arm) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(arm), levels = c("None", comparison)))

      if (length(unique(df_primary$CNV)) == 2) {
        surv_obj <- Surv(df_primary$OS_MONTHS, df_primary$OS)
        p_primary <- tryCatch({
          1 - pchisq(survdiff(surv_obj ~ CNV, data = df_primary)$chisq, df = 1)
        }, error = function(e) NA)
      } else {
        p_primary <- NA
      }

      # METASTASIS
      df_met <- tumor_df %>%
        filter(SAMPLE_TYPE == "Metastasis", !!sym(arm) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(arm), levels = c("None", comparison)))

      if (length(unique(df_met$CNV)) == 2) {
        surv_obj <- Surv(df_met$OS_MONTHS, df_met$OS)
        p_met <- tryCatch({
          1 - pchisq(survdiff(surv_obj ~ CNV, data = df_met)$chisq, df = 1)
        }, error = function(e) NA)
      } else {
        p_met <- NA
      }

      primary_met_results <- rbind(primary_met_results, data.frame(
        CANCER_TYPE = tumor,
        Arm = arm,
        Comparison = comparison,
        P_primary = signif(p_primary, 3),
        N_primary = nrow(df_primary),
        P_metastasis = signif(p_met, 3),
        N_metastasis = nrow(df_met)
      ))
    }
  }
}

# View the output
print(primary_met_results, n = Inf)
# write.csv(primary_met_results, "Primary_vs_Metastasis_Survival_Comparison.csv", row.names = FALSE)
```

```{r}
# Filter to only rows where either p_primary or p_metastasis is < 0.1
filtered_results <- primary_met_results %>%
  filter((!is.na(P_primary) & P_primary < 0.1) | (!is.na(P_metastasis) & P_metastasis < 0.1))

# Print results
print(filtered_results, n = Inf)
# write.csv(filtered_results, "Primary_vs_Metastasis_Survival_P0.1.csv", row.names = FALSE)
```


Visual comparision 
```{r}
library(tidyverse)

# Filtered input (from primary_met_results)
plot_data <- primary_met_results %>%
  filter(!is.na(P_primary) | !is.na(P_metastasis)) %>%
  mutate(
    logP_primary = -log10(P_primary),
    logP_metastasis = -log10(P_metastasis),
    Comparison_Label = paste(Arm, Comparison)
  ) %>%
  filter(!is.na(logP_primary) | !is.na(logP_metastasis)) %>%
  pivot_longer(
    cols = c(logP_primary, logP_metastasis),
    names_to = "Group", values_to = "logP"
  ) %>%
  mutate(
    Group = recode(Group,
                   logP_primary = "Primary",
                   logP_metastasis = "Metastasis")
  )

# Custom color palette
custom_colors <- c("Primary" = "#2ca02c",     # green
                   "Metastasis" = "#9467bd")  # purple

# Loop through and save one plot per cancer type
for (tumor in unique(plot_data$CANCER_TYPE)) {
  p <- ggplot(filter(plot_data, CANCER_TYPE == tumor),
              aes(x = Comparison_Label, y = logP, fill = Group)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
    geom_hline(yintercept = -log10(0.1), linetype = "dashed", color = "red") +
    labs(
      title = paste0("-log10(p) Survival Comparison — ", tumor),
      x = "CNV Arm + Event",
      y = "-log10(p)",
      fill = "Sample Type"
    ) +
    scale_fill_manual(values = custom_colors) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, size = 8),
      plot.title = element_text(face = "bold", size = 14)
    )

  # Save plot or print
  print(p)
  ggsave(paste0("PvsM_survival_comparison/", gsub(" ", "_", tumor), ".pdf"), p, width = 10, height = 5)
}
```


## NSLC 
```{r}
unique(clin_smp$CANCER_TYPE_DETAILED)
```


```{r}
library(tidyverse)
library(survival)

arm_cols <- grep("^Arm", names(df), value = TRUE)
subtype_results <- data.frame()

# Filter to NSCLC samples
nsclc_df <- df %>% filter(CANCER_TYPE == "Non-Small Cell Lung Cancer")

# Loop over arms and Gain/Loss
for (arm in arm_cols) {
  for (comparison in c("Gain", "Loss")) {
    
    for (subtype in c("Lung Adenocarcinoma", "Lung Squamous Cell Carcinoma")) {
      
      subset <- nsclc_df %>%
        filter(CANCER_TYPE_DETAILED == subtype,
               !!sym(arm) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(arm), levels = c("None", comparison)))
      
      if (length(unique(subset$CNV)) == 2) {
        surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
        p_val <- tryCatch({
          1 - pchisq(survdiff(surv_obj ~ CNV, data = subset)$chisq, df = 1)
        }, error = function(e) NA)
      } else {
        p_val <- NA
      }
      
      subtype_results <- rbind(subtype_results, data.frame(
        MainCancer = "Non-Small Cell Lung Cancer",
        Subtype = subtype,
        Arm = arm,
        Comparison = comparison,
        P = signif(p_val, 3),
        N = nrow(subset)
      ))
    }
  }
}

print(subtype_results, n = Inf)
# Save results
# write.csv(subtype_results, "NSCLC_Subtype_Comparison.csv", row.names = FALSE)

# Save to CSV
#write.csv(subtype_results, "Subtype_Survival_Pvalues.csv", row.names = FALSE)
#write.csv(sig_results, "Subtype_Survival_P_lt_0.1.csv", row.names = FALSE)
```

```{r}
# Filter by p < 0.1
subtype_results <- subtype_results %>%
  filter(!is.na(P) & P < 0.1)

# Print instead of saving to CSV
print(subtype_results, n = Inf)
```
- only lung adenocarcinoma KM curves 
```{r}
# Output directory
output_dir <- "KM_plots_PD1_0.3_LUAD"
dir.create(output_dir, showWarnings = FALSE)

# Initialize table
km_sig_hits <- data.frame()

# Subset to LUAD within NSCLC
tumor_data <- df %>%
  filter(CANCER_TYPE == "Non-Small Cell Lung Cancer",
         CANCER_TYPE_DETAILED == "Lung Adenocarcinoma",
         DRUG_TYPE == "PD-1/PDL-1")

for (arm in grep("^Arm", names(tumor_data), value = TRUE)) {
  for (comparison in c("Gain", "Loss")) {
    subset <- tumor_data %>%
      filter(!!sym(arm) %in% c("None", comparison)) %>%
      mutate(CNVstatus = factor(!!sym(arm), levels = c(comparison, "None")))

    if (nrow(subset) >= 10 && length(unique(subset$CNVstatus)) == 2) {
      surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
      surv_diff <- survdiff(surv_obj ~ CNVstatus, data = subset)
      km_p <- 1 - pchisq(surv_diff$chisq, df = 1)

      if (km_p < 0.1) {
        km_sig_hits <- rbind(km_sig_hits, data.frame(
          CANCER_TYPE = "NSCLC",
          Subtype = "Lung Adenocarcinoma",
          Arm = arm,
          Comparison = comparison,
          KMP = signif(km_p, 3),
          N = nrow(subset)
        ))

        # Plotting
        fit <- survfit(surv_obj ~ CNVstatus, data = subset)

        current_palette <- if (comparison == "Gain") {
          c("#EA6A6A", "black")
        } else {
          c("#4F6DB8", "black")
        }

        p <- ggsurvplot(fit, data = subset,
                        pval = TRUE,
                        risk.table = TRUE,
                        conf.int = FALSE,
                        palette = current_palette,
                        title = paste("LUAD", arm, comparison, "vs None – PD-1/PDL-1"),
                        xlab = "Survival Time (months)",
                        ylab = "Survival probability",
                        risk.table.height = 0.25,
                        break.time.by = 20,
                        ggtheme = theme_classic(base_size = 14),
                        linetype = "solid",
                        size = 1.4)

        combined_plot <- arrangeGrob(p$plot, p$table, ncol = 1, heights = c(3, 1))
        ggsave(filename = paste0(output_dir, "/", arm, "_", comparison, "_LUAD_PD1.pdf"),
               plot = combined_plot, width = 7, height = 7)
      }
    }
  }
}

# Print only significant hits
print(km_sig_hits, n = Inf)
```



## Glioma 
```{r}
library(tidyverse)
library(survival)

# Identify CNV arm columns
arm_cols <- grep("^Arm", names(df), value = TRUE)

# Subtypes to include (from CANCER_TYPE_DETAILED column)
glioma_subtypes <- c(
  "Glioblastoma Multiforme",
  "High-Grade Glioma, NOS",
  "Low-Grade Glioma, NOS",
  "Gliosarcoma",
  "Glioblastoma",
  "Astrocytoma",
  "Anaplastic Astrocytoma",
  "Anaplastic Oligodendroglioma",
  "Oligodendroglioma"
)

# Filter for Glioma only
glioma_df <- df %>% filter(CANCER_TYPE == "Glioma")

# Output container
subtype_results <- data.frame()

# Loop over CNV arms and Gain/Loss comparisons
for (arm in arm_cols) {
  for (comparison in c("Gain", "Loss")) {
    for (subtype in glioma_subtypes) {
      
      subset <- glioma_df %>%
        filter(CANCER_TYPE_DETAILED == subtype,
               !!sym(arm) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(arm), levels = c("None", comparison)))
      
      if (length(unique(subset$CNV)) == 2 && nrow(subset) >= 10) {
        surv_obj <- Surv(subset$OS_MONTHS, subset$OS)
        p_val <- tryCatch({
          1 - pchisq(survdiff(surv_obj ~ CNV, data = subset)$chisq, df = 1)
        }, error = function(e) NA)
      } else {
        p_val <- NA
      }
      
      subtype_results <- rbind(subtype_results, data.frame(
        MainCancer = "Glioma",
        Subtype = subtype,
        Arm = arm,
        Comparison = comparison,
        P = signif(p_val, 3),
        N = nrow(subset)
      ))
    }
  }
}

# View all results
print(subtype_results, n = Inf)
```

```{r}
# Filter for significant results only
subtype_results_sig <- subtype_results %>%
  filter(!is.na(P) & P < 0.1)

# Print all significant results
print(subtype_results_sig)
```

- high grade vs low grade
