---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(dplyr)

# Folder path
folder_path <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/TCGA_Deconvolutions_CNV_IS/Arm/Outputs"

# Match continuous regression files
files <- list.files(
  path = folder_path,
  pattern = "continuous0.2_regression.csv",  # looser match
  full.names = TRUE
)

# Print what you found
print(basename(files))

# Load all matching files
all_data <- lapply(files, function(file) {
  tryCatch({
    df <- read_csv(file)
    df$TumorType <- sub("_is0.3.*", "", basename(file))  # Extract tumor type
    return(df)
  }, error = function(e) {
    message("Failed to read: ", file)
    return(NULL)
  })
}) %>% bind_rows()

# Check result
colnames(all_data)
head(all_data)
```
# Filter for selected tumor types
```{r}
tumor_types_in_figure <- c(
  "UCEC.MSI", "UCEC.MSS", "OV", "CESC", "LUSC", 
  "HNSC.HPVneg", "ESCA.SC", "BLCA", "BRCA", "BRCA.neg", "BRCA.pos","PRAD",
  "LUAD", "LIHC", "COADREAD.MSI", "COADREAD.MSS", "ESCA.AD",  "STAD", "PAAD", 
  "KICH", "KIRC", "KIRP", "LGG", "GBM", "SKCM", "UVM", "ACC","MESO", "PCPG", "UCS", "SARC", "TGCT"
)

# Example: Filter your main dataset 'all_data' 
# so that it only contains the tumor types from the figure.
filtered_data <- all_data %>%
  filter(TumorType %in% tumor_types_in_figure)
```


# Fig2
## SKCM
```{r}
SKCM_targets <- data.frame(
  Arm = c("1q", "4q", "6p", "6q", "9p"),
  Event = c("Gain", "Loss", "Gain", "Loss", "Loss")
)

# Filter the data based on SKCM_targets
SKCM <- all_data %>%
  filter(TumorType == "SKCM") %>%
  semi_join(SKCM_targets, by = c("Arm", "Event")) %>%
  arrange(Arm, P_val)

# Print results
print(SKCM)
```

## LUAD
```{r}
library(dplyr)

# Define arm-event pairs of interest
luad_targets <- data.frame(
  Arm = c("1q", "5q", "6q", "9p"),
  Event = c("Gain", "Loss", "Loss", "Loss")
)

# Filter the data
luad <- all_data %>%
  filter(
    TumorType == "LUAD" & (
      (Arm == "1q" & Event == "Gain") |
      (Arm == "5q" & Event == "Loss") |
      (Arm == "6q" & Event == "Loss") |
      (Arm == "9p" & Event == "Loss") |
      (Arm == "15q" & Event == "Loss"))
  ) %>%
  arrange(Arm, P_val)

# Print results
print(luad)
```

```{r}
# Load the RDS
tcga <- readRDS("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/scores/TCGA_working_tables.rds")

# Extract LUAD arm CNV matrix
luad_arm <- tcga[["LUAD_ArmCNV"]]

# Add rownames as a column BEFORE reordering
luad_arm$Sample_Barcode <- rownames(luad_arm)

# Move Sample_Barcode to the first column
luad_arm <- luad_arm[, c("Sample_Barcode", setdiff(names(luad_arm), "Sample_Barcode"))]

# Save CSV with correct sample identifiers
write.csv(luad_arm, "LUAD_ArmCNV.csv", row.names = FALSE)
```

## BLCA
```{r}
BLCA_targets <- data.frame(
  Arm = c("9p", "11p", "17q"),
  Event = c("Loss", "Loss", "Gain")
)

# Filter the data based on SKCM_targets
BLCA <- all_data %>%
  filter(TumorType == "BLCA") %>%
  semi_join(BLCA_targets, by = c("Arm", "Event")) %>%
  arrange(Arm, P_val)

# Print results
print(BLCA)
```

## PAAD
```{r}
PAAD_targets <- data.frame(
  Arm = c("1q", "6q", "9p"),
  Event = c("Gain", "Loss", "Loss")
)

# Filter the data based on SKCM_targets
PAAD <- all_data %>%
  filter(TumorType == "PAAD") %>%
  semi_join(PAAD_targets, by = c("Arm", "Event")) %>%
  arrange(Arm, P_val)

# Print results
print(PAAD)
```

## KIRC
```{r}
kirc_9p_gain_lowIS <- filtered_cytoband_data %>%
  filter(
    TumorType == "KIRC",
    Chromosome == "Chr9",
    Arm == "p",
    Event == "Gain",
    Z_val < 0,  # lower IS association
    !is.na(Z_val)
  ) %>%
  arrange(P_val)  # optional: sort by significance

print(kirc_9p_gain_lowIS)
```

# Get the desired arms
- The Coefficients represent the effect size of CNV (e.g., arm loss) on cytotoxic immune score (IS).
- A negative Coefficient = CNV reduces IS → lower cytotoxic immune score

```{r}
# Define arms of interest
target_arms <- c("9p", "1p", "5q", "4q")

# Filter for loss events on those arms
results <- filtered_data %>%
  filter(Arm %in% target_arms, Event == "Loss") %>%
  select(TumorType, Arm, Event, Coefficients, Z_val, P_val)

# Preview the results
print(results)
```


```{r}
# For each arm, get the top 20 rows (or however many you want) with p-value < 0.05,
# ordered by p-value (i.e., the most significant ones first)
top20_significant <- filtered_data %>%
  filter(Arm %in% target_arms, Event == "Loss", P_val < 0.2) %>%
  group_by(Arm) %>%
  arrange(P_val, .by_group = TRUE) %>%
  slice_head(n = 10) %>%
  select(Arm, TumorType, Coefficients, Z_val, P_val)

# Check final result
print(top20_significant)

# Save to CSV
# write.csv(top20_significant, "Top20_Significant_by_Pval.csv", row.names = FALSE)
```

```{r}
 # Filter the data for 5q loss and the selected tumor types, then order by p-value
filtered_5q_loss <- filtered_data %>%
  filter(Arm == "5q", Event == "Loss") %>%
  arrange(P_val)

# Print the result
print(filtered_5q_loss)
```

```{r}
 # Filter the data for 4q loss and the selected tumor types, then order by p-value
filtered_4q_loss <- filtered_data %>%
  filter(Arm == "4q", Event == "Loss") %>%
  arrange(P_val)

# Print the result
print(filtered_4q_loss)
```

```{r}
 # Filter the data for 4p loss and the selected tumor types, then order by p-value
filtered_1p_loss <- filtered_data %>%
  filter(Arm == "1p", Event == "Loss") %>%
  arrange(P_val)

# Print the result
print(filtered_1p_loss)
```

```{r}
top10_lowIS <- all_data %>%
  filter(
    Arm %in% target_arms,
    Event == "Loss",
    P_val < 0.05
  ) %>%
  group_by(Arm) %>%
  arrange(Coefficients, .by_group = TRUE) %>%  # sort by IS suppression
  slice_head(n = 10) %>%
  ungroup() %>%
  select(Arm, TumorType, Coefficients, Z_val, P_val)

# Print full result (not truncated)
print(top10_lowIS, n = Inf)


```

```{r}
# Define target arms for gains
target_gain_arms <- c("1q", "6p", "8p", "17q")

# Filter for arm-level GAIN events with significant positive effect on IS
top_gain <- filtered_data %>%
  filter(
    Arm %in% target_gain_arms,
    Event == "Gain",
    P_val<0.2
  ) %>%
  group_by(Arm) %>%
  arrange(P_val, .by_group = TRUE) %>%   
  select(Arm, TumorType, Coefficients, Z_val, P_val)

# Print all results
print(top_gain, n = Inf)
```

```{r}
 # Filter the data for 4p loss and the selected tumor types, then order by p-value
filtered_8p_gain <- filtered_data %>%
  filter(Arm == "8p", Event == "Gain") %>%
  arrange(P_val)

# Print the result
print(filtered_8p_gain)
```

# Cytotoxic scores 
```{r}
library(readr)
library(dplyr)

# Folder path
folder_path <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/bulk seq/scores"

# Match continuous regression files
files <- list.files(
  path = folder_path,
  pattern = "regression_results_continous.csv",  # looser match
  full.names = TRUE
)

# Print what you found
print(basename(files))

# Load all matching files
all_data <- lapply(files, function(file) {
  tryCatch({
    df <- read_csv(file)
    df$TumorType <- sub("_is0.3.*", "", basename(file))  # Extract tumor type
    return(df)
  }, error = function(e) {
    message("Failed to read: ", file)
    return(NULL)
  })
}) %>% bind_rows()

# Combine CD8_T_cells and NK_cells to form a cytotoxic immune score
selected_columns <- grep("^(TumorType|CD8_T_cells\\.|NK_cells\\.)", colnames(all_data), value = TRUE)
cytotoxic_data <- all_data[, selected_columns]

# Check result
colnames(cytotoxic_data)
head(cytotoxic_data)
```

```{r}
# Define arms of interest
target_arms <- c("9p", "1p", "5q", "4q")

# Filter for loss events on those arms
# For each arm, get top 7 lowest Coefficient (i.e., lowest IS)
top10_lowIS <- cytotoxic_data %>%
  filter(Arm %in% target_arms, Event == "Loss") %>%
  group_by(Arm) %>%
  arrange(Cytotoxic_Coefficients) %>%
  slice_head(n = 20) %>%
  select(Arm, TumorType, Cytotoxic_Coefficients, Cytotoxic_Z_val, Cytotoxic_P_val)

print(top10_lowIS)
```




# Cytobands
```{r}
library(readr)
library(dplyr)

# === Cytoband-level regression folder ===
folder_path <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/TCGA_Deconvolutions_CNV_IS/Cytoband/Outputs"

# Match continuous regression files (same pattern)
files <- list.files(
  path = folder_path,
  pattern = "continuous0.2_regression.csv",  # adjust pattern if needed
  full.names = TRUE
)

# Print what was found
print(basename(files))

# Load all cytoband regression files
all_cytoband_data <- lapply(files, function(file) {
  tryCatch({
    df <- read_csv(file)
    df$TumorType <- sub("_is0.3.*", "", basename(file))  # Extract tumor type
    return(df)
  }, error = function(e) {
    message("Failed to read: ", file)
    return(NULL)
  })
}) %>% bind_rows()

# View column names and preview data
colnames(all_cytoband_data)
head(all_cytoband_data)
```

```{r}

# Filter to only specific tumor types 
tumor_types_in_figure <- c(
  "UCEC.MSI", "UCEC.MSS", "OV", "CESC", "LUSC", 
  "HNSC.HPVneg", "ESCA.SC", "BLCA", "BRCA", "BRCA.neg", "BRCA.pos","PRAD",
  "LUAD", "LIHC", "COADREAD.MSI", "COADREAD.MSS", "ESCA.AD",  "STAD", "PAAD", 
  "KICH", "KIRC", "KIRP", "LGG", "GBM", "SKCM", "UVM", "ACC","MESO", "PCPG", "UCS", "SARC", "TGCT"
)

# Example: Filter your main dataset 'all_data' 
# so that it only contains the tumor types from the figure.
filtered_cytoband_data <- all_cytoband_data %>%
  filter(TumorType %in% tumor_types_in_figure)

# Check result
colnames(all_cytoband_data)
colnames(filtered_cytoband_data)
head(filtered_cytoband_data)
#write.csv(filtered_cytoband_data, "Selected_Cytoband_Events.csv", row.names = FALSE)
```

```{r}
# Print all cytobands starting with 6p or 8p and are Gain events
gain_6p_8p <- filtered_cytoband_data %>%
  filter(
    Event == "Gain",
    P_val < 0.05,
    grepl("^6p", Cytoband) | grepl("^8p", Cytoband)
  ) %>%
  arrange(Cytoband, TumorType)

# View result
print(gain_6p_8p, n = Inf)
```

```{r}
# Print all cytobands starting with 6p or 8p and are Gain events
loss_cytoband <- filtered_cytoband_data %>%
  filter(
    Event == "Loss",
    P_val < 0.05,
    grepl("^18q", Cytoband)  
  ) %>%
  arrange(Cytoband, TumorType)

# View result
print(loss_cytoband, n = Inf)
```

## Print out the table containing all cytobands and their tumor types 
```{r}
library(dplyr)
library(tidyr)

wide_table <- filtered_cytoband_data %>%
  # Keep only the relevant columns
  select(Cytoband, TumorType, Z_val, P_val) %>%
  # Pivot to wide format so each Cytoband becomes a row and each TumorType provides its Z and P values
  pivot_wider(
    id_cols = Cytoband,
    names_from = TumorType,
    values_from = c(Z_val, P_val),
    names_glue = "{TumorType}-{.value}",
    # In case there are any duplicates, use the first value
    values_fn = list(Z_val = ~ .[1], P_val = ~ .[1])
  ) %>%
  # Rename 
  rename_with(~ gsub("Z_val", "Z", .x)) %>%
  rename_with(~ gsub("P_val", "p", .x))

# Print the final wide-format table
print(wide_table)
write.csv(wide_table, "Significant_Cytobands.csv", row.names = FALSE)
```




## Group all cytobands together 
```{r}
# Step 1: Create a grouping column for cytobands (e.g., "9p21" from "9p21.1", "9p21.2", etc.)
aggregated_data <- filtered_cytoband_data %>%
  mutate(Cytoband_group = sub("\\..*$", "", Cytoband)) %>%
  # Step 2: For each TumorType and Cytoband_group, summarize by taking the minimum p-value 
  # and the corresponding Z value (you can adjust the summarization as needed)
  group_by(TumorType, Cytoband_group) %>%
  summarise(
    min_P_val = min(P_val, na.rm = TRUE),
    # Get the Z_val corresponding to the minimum p-value:
    corresponding_Z = Z_val[which.min(P_val)],
    count = n(),  # optional: count how many sub-cytobands were aggregated
    .groups = "drop"
  )

# Step 3: For each tumor type, select the cytoband group with the lowest p-value
most_significant <- aggregated_data %>%
  group_by(TumorType) %>%
  slice_min(min_P_val, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(TumorType, min_P_val)

# Print the result
print(most_significant)
```

## 9p loss
```{r}
filtered9p_data <- filtered_cytoband_data %>%
  # Filter to only chromosome 9p loss events
  filter(Chromosome == "Chr9", Arm == "p", Event == "Loss") %>%
  # Create a new column that removes the decimal part from Cytoband (e.g. "9p21.1" -> "9p21")
  mutate(Cytoband_group = sub("\\..*$", "", Cytoband)) %>%
  # Group by TumorType and the simplified Cytoband
  group_by(TumorType, Cytoband_group) %>%
  summarise(
    mean_P_val = mean(P_val, na.rm = TRUE),
    mean_Z     = mean(Z_val, na.rm = TRUE),
    count      = n(),  # count how many sub-cytobands contributed
    .groups = "drop"
  ) %>%
  # Filter for rows with mean_P_val less than 0.1
  filter(mean_P_val < 0.05) %>%
  # Arrange first by Cytoband_group, then by mean_P_val (lowest p-value = most significant)
  arrange(Cytoband_group, mean_P_val)

print(filtered9p_data)

```

```{r}
# print out significant cytobands
# Define the critical cytobands
library(stringr)

critical_cytobands <- c("9p21", "9p13", "9p24")

critical_data <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr9",
    Arm == "p",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(P_val)  # Order by p-value

print(critical_data)
```

p values combined fisher test
```{r}
library(dplyr)

data_HNSC <- filtered_cytoband_data %>%
  filter(Chromosome == "Chr9", Arm == "p", Event == "Gain", TumorType == "HNSC.HPVneg") %>%
  arrange(Cytoband)

print(data_HNSC)
```

```{r}
library(dplyr)

data_BLCA <- filtered_cytoband_data %>%
  filter(Chromosome == "Chr9", Arm == "p", Event == "Gain", TumorType == "BLCA") %>%
  arrange(Cytoband)

print(data_BLCA)
```

## 1q gain 
```{r}
oneqgain <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr1",
    Arm == "q",
    Event == "Gain",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(oneqgain)
```


## 1p loss 
```{r}
oneploss <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr1",
    Arm == "p",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(oneploss)
```

## 4q loss
```{r}
fourqloss <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr4",
    Arm == "q",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(fourqloss)
```

## 5q loss
```{r}
fiveqloss <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr5",
    Arm == "q",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(fiveqloss)
```

## 18q loss
```{r}
q18loss <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr18",
    Arm == "q",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(q18loss)
```

## 17q gain
```{r}
q17gain <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr17",
    Arm == "q",
    Event == "Gain",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(q17gain)
```

## 4q loss
```{r}
q4loss <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr4",
    Arm == "q",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(Cytoband, TumorType)  # Order by p-value

print(q4loss)
```

## 9p loss
```{r}
p9loss <- filtered_cytoband_data %>%
  filter(
    Chromosome == "Chr9",
    Arm == "p",
    Event == "Loss",
    P_val < 0.05,  # Only include rows with P_val < 0.1
    # Check if Cytoband starts with one of the critical values
    #str_detect(Cytoband, paste0("^(", paste(critical_cytobands, collapse = "|"), ")"))
  ) %>%
  arrange(P_val)  # Order by p-value

print(p9loss)
```


# Immune hot
```{r}
# Define acrocentric chromosome arms
acrocentric_arms <- c("13q", "14q", "15q", "21q", "22q")

# Filter for LOSS events on acrocentric arms that are IMMUNE HOT (Z > 0, p < 0.05)
immune_hot_losses <- filtered_data %>%
  filter(
    Arm %in% acrocentric_arms,
    Event == "Loss",
    Z_val > 0
  ) %>%
  arrange(Arm, TumorType, P_val)

# Print result
print(immune_hot_losses, n = Inf)
```

