---
title: "R Notebook"
output: html_notebook
---

Make sure arm-level prediction 
- correlation between arm status & global AS 
  - compute Arm value or binary gain/loss with global AS 
- cross-task comparison between 2 models 

# HNSC HPVneg
```{r}
library(tidyverse)
library(tibble)

tcga <- readRDS("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/scores/TCGA_working_tables.rds")
names(tcga) # confirms what is inside 

tcga_hnsc <- tcga[["HNSC.HPVneg_ArmCNV"]]  # Use HPV– subset
colnames(tcga[["HNSC.HPVneg_ArmCNV"]])[1:30] # see column names inside tcga_hnsc

# Identify arm columns safely
arm_cols_tcga <- grep("^Arm", colnames(tcga_hnsc), value = TRUE)
print(length(arm_cols_tcga))
print(head(arm_cols_tcga))

stopifnot(length(arm_cols_tcga) > 0)

# Subset arms only
tcga_arms_only <- tcga_hnsc[, arm_cols_tcga, drop = FALSE]

# Compute global aneuploidy score = #arms with |value| > 0.2 (based on manuscript )
tcga_global <- data.frame(
  Patient_ID = rownames(tcga_hnsc),
  Tumor = "HNSC",
  Global_AS = rowSums(abs(tcga_arms_only) > 0.2, na.rm = TRUE)
)

summary(tcga_global$Global_AS)
head(tcga_global)
```

```{r}
# split AS by median for binarization for the model 
median_AS <- median(tcga_global$Global_AS, na.rm = TRUE)

# High_AS = 1 if Global_AS ≥ median, else 0
tcga_global$High_Global_AS <- ifelse(
  tcga_global$Global_AS > median_AS,
  1,
  0
)
```

sanity check 

```{r}
table(tcga_global$High_Global_AS)
prop.table(table(tcga_global$High_Global_AS))
summary(tcga_global$Global_AS)
```

```{r}
# confirm rownames are patient/sample IDs
head(rownames(tcga_hnsc), 5)
head(tcga_global$Patient_ID, 5)
```

save the label csv
```{r}
out_csv <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/image_analysis/HNSC_HPVneg_globalAS_labels.csv"
write_csv(tcga_global, out_csv)
```



```{r}
cptac <- read.csv("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/cptac_pancan/Validation/CPTAC_PanCan_Arm_Level_CopyNumber_NORMALIZED.csv", 
                  row.names = 1, check.names = FALSE)
head(cptac)
unique(cptac$TumorType)
```

```{r}

# Filter only HNSCC rows
cptac_hnsc <- cptac[cptac$TumorType == "HNSCC", ]

# Convert rownames (patient IDs) to a column
cptac_hnsc <- rownames_to_column(cptac_hnsc, "Patient_ID")
cptac_hnsc

# Generate label file: binary 1q gain
cptac_labels <- cptac_hnsc %>%
  mutate(
    Tumor = "HNSC",
    Arm_1q_Gain = ifelse(Arm_1q > 0.2, 1, 0)
  ) %>%
  select(Patient_ID, Tumor, Arm_1q_Gain)

# Check it
cptac_labels

```


```{r}
# ==== Combine CPTAC + TCGA ====
# Combine TCGA and CPTAC
combined_labels <- bind_rows(cptac_labels, tcga_labels)

# Preview
combined_labels

# Save to file
write.csv(combined_labels, "combined_HNSC_Arm1qGain_labels.csv", row.names = FALSE)
```

# Tile 
Check match with tile 
```{r}
library(tidyverse)

# Load
setwd("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/image_analysis")
tiles <- read.csv("all_tiles.csv")
labels <- read.csv("combined_HNSC_Arm1qGain_labels.csv")


# Check match
intersect_ids <- intersect(unique(tiles$Patient_ID), labels$Patient_ID)

length(intersect_ids)       # number of matched patients
length(unique(tiles$Patient_ID))  # total in all_tiles
length(labels$Patient_ID)   # total in label file
```

Filter to only matched

## Match 
```{r}
tiles_filtered <- tiles %>% filter(Patient_ID %in% labels$Patient_ID)
write.csv(tiles_filtered, "all_tiles_1qgain_filtered.csv", row.names = FALSE)
```


## aneuplody patient
```{r}
tcga_hnsc <- tcga[["HNSC.HPVneg_ArmCNV"]]

# Compute aneuploidy
tcga_hnsc$Aneuploidy_Score <- apply(tcga_hnsc, 1, function(x) sum(abs(x) > 0.2, na.rm = TRUE))
tcga_hnsc$Patient_ID <- paste0(rownames(tcga_hnsc), "-tumor")

# Subset 9p loss (< -0.2) vs neutral (≥ -0.2)
tcga_9p_loss <- tcga_hnsc[tcga_hnsc$Arm9p < -0.2, ]
tcga_9p_neutral <- tcga_hnsc[tcga_hnsc$Arm9p >= -0.2, ]

# Find most aneuploid in each group
tcga_top_9p_loss <- tcga_9p_loss[which.max(tcga_9p_loss$Aneuploidy_Score), "Patient_ID"]
tcga_top_9p_neutral <- tcga_9p_neutral[which.max(tcga_9p_neutral$Aneuploidy_Score), "Patient_ID"]

cat("TCGA HNSC.HPVneg most aneuploid 9p loss patient:", tcga_top_9p_loss, "\n")
cat("TCGA HNSC.HPVneg most aneuploid 9p neutral patient:", tcga_top_9p_neutral, "\n")
```

```{r}
# Top 3 most aneuploid TCGA 9p loss patients
tcga_top3_9p_loss <- tcga_9p_loss[order(-tcga_9p_loss$Aneuploidy_Score), ]
top3_loss_ids <- head(tcga_top3_9p_loss$Patient_ID, 3)

cat("Top 3 TCGA HNSC.HPVneg 9p loss patients:\n")
print(top3_loss_ids)
```

```{r}
# Load CPTAC
cptac <- read.csv("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/cptac_pancan/Validation/CPTAC_PanCan_Arm_Level_CopyNumber_NORMALIZED.csv", 
                  row.names = 1, check.names = FALSE)

# Filter HNSCC
library(tibble)
cptac_hnsc <- cptac[cptac$TumorType == "HNSCC", ]
cptac_hnsc <- rownames_to_column(cptac_hnsc, "Patient_ID")

# Compute aneuploidy score
cptac_hnsc$Aneuploidy_Score <- apply(
  cptac_hnsc[, !(colnames(cptac_hnsc) %in% c("TumorType", "Patient_ID"))],
  1,
  function(x) sum(abs(x) > 0.2, na.rm = TRUE)
)

# Subset by 9p status
cptac_9p_loss <- cptac_hnsc[cptac_hnsc$Arm_9p < -0.2, ]
cptac_9p_neutral <- cptac_hnsc[cptac_hnsc$Arm_9p >= -0.2, ]

# Most aneuploid in each group
cptac_top_9p_loss <- cptac_9p_loss[which.max(cptac_9p_loss$Aneuploidy_Score), "Patient_ID"]
cptac_top_9p_neutral <- cptac_9p_neutral[which.max(cptac_9p_neutral$Aneuploidy_Score), "Patient_ID"]

cat("CPTAC HNSCC most aneuploid 9p loss patient:", cptac_top_9p_loss, "\n")
cat("CPTAC HNSCC most aneuploid 9p neutral patient:", cptac_top_9p_neutral, "\n")
```

# HNSC 9p loss
```{r}
# --- TCGA ---
tcga <- readRDS("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/scores/TCGA_working_tables.rds")
tcga_hnsc <- tcga[["HNSC.HPVneg_ArmCNV"]]

# Create TCGA 9p loss label file
tcga_labels_9p <- data.frame(
  Patient_ID = paste0(rownames(tcga_hnsc), "-tumor"),
  Tumor = "HNSC",
  Arm_9p_Loss = ifelse(tcga_hnsc$Arm9p < -0.2, 1, 0)
)
```

```{r}
# --- CPTAC ---
cptac <- read.csv("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/cptac_pancan/Validation/CPTAC_PanCan_Arm_Level_CopyNumber_NORMALIZED.csv", 
                  row.names = 1, check.names = FALSE)

# Filter HNSCC only
cptac_hnsc <- cptac[cptac$TumorType == "HNSCC", ]
cptac_hnsc <- tibble::rownames_to_column(cptac_hnsc, "Patient_ID")

# Create CPTAC 9p loss labels
cptac_labels_9p <- cptac_hnsc %>%
  mutate(
    Tumor = "HNSC",
    Arm_9p_Loss = ifelse(Arm_9p < -0.2, 1, 0)
  ) %>%
  select(Patient_ID, Tumor, Arm_9p_Loss)
```

```{r}
# --- Combine TCGA + CPTAC ---
combined_labels_9p <- bind_rows(tcga_labels_9p, cptac_labels_9p)

# Save to file
write.csv(combined_labels_9p, "combined_HNSC_Arm9pLoss_labels.csv", row.names = FALSE)
```

```{r}
tiles <- read.csv("all_tiles.csv")
labels_9p <- read.csv("combined_HNSC_Arm9pLoss_labels.csv")

# Match Patient IDs
intersect_ids_9p <- intersect(unique(tiles$Patient_ID), labels_9p$Patient_ID)

length(intersect_ids_9p)
length(unique(tiles$Patient_ID))
length(labels_9p$Patient_ID)
```

```{r}
tiles_filtered_9p <- tiles %>% filter(Patient_ID %in% labels_9p$Patient_ID)
write.csv(tiles_filtered_9p, "all_tiles_9ploss_filtered.csv", row.names = FALSE)
```

```{r}
# check match between lisa and mine
# Load both CSV files
df1 <- read.csv("HNSC.HPVneg/arm9p_loss_labels.csv")
df2 <- read.csv("HNSC.HPVneg/combined_HNSC_Arm9pLoss_labels.csv")

# Check dimensions
cat("df1 rows:", nrow(df1), "\n")
cat("df2 rows:", nrow(df2), "\n")

# Print column names
cat("df1 columns:\n"); print(colnames(df1))
cat("df2 columns:\n"); print(colnames(df2))

# Check how many shared Patient_IDs
shared_ids <- intersect(df1$Patient_ID, df2$Patient_ID)
cat("Number of shared Patient_IDs:", length(shared_ids), "\n")

# Show example of Patient_IDs only in df1
only_in_df1 <- setdiff(df1$Patient_ID, df2$Patient_ID)
cat("Patient_IDs only in df1 (first 10):\n")
print(head(only_in_df1, 10))

# Show example of Patient_IDs only in df2
only_in_df2 <- setdiff(df2$Patient_ID, df1$Patient_ID)
cat("Patient_IDs only in df2 (first 10):\n")
print(head(only_in_df2, 10))
```
## TCGA + CPTAC separately 
```{r}
# --- TCGA only ---
tcga <- readRDS("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/scores/TCGA_working_tables.rds")
tcga_hnsc <- tcga[["HNSC.HPVneg_ArmCNV"]]

# 9p Loss label creation
tcga_labels_9p <- data.frame(
  Patient_ID = paste0(rownames(tcga_hnsc), "-tumor"),
  Tumor = "HNSC",
  Arm_9p_Loss = ifelse(tcga_hnsc$Arm9p < -0.2, 1, 0)676
)

write.csv(tcga_labels_9p, "tcga_HNSC_Arm9pLoss_labels.csv", row.names = FALSE)

# --- CPTAC only ---
cptac <- read.csv("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/cptac_pancan/Validation/CPTAC_PanCan_Arm_Level_CopyNumber_NORMALIZED.csv", row.names = 1, check.names = FALSE)

# Filter to HNSCC and move rownames to column
cptac_hnsc <- cptac[cptac$TumorType == "HNSCC", ] %>% rownames_to_column("Patient_ID")

# 9p Loss label creation
cptac_labels_9p <- cptac_hnsc %>%
  mutate(
    Tumor = "HNSC",
    Arm_9p_Loss = ifelse(Arm_9p < -0.2, 1, 0)
  ) %>%
  select(Patient_ID, Tumor, Arm_9p_Loss)

write.csv(cptac_labels_9p, "cptac_HNSC_Arm9pLoss_labels.csv", row.names = FALSE)
```


# UCEC.MSI 
```{r}
library(tidyverse)

tcga <- readRDS("/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/TCGA/scores/TCGA_working_tables.rds")
names(tcga)
```


```{r}
# Pull UCEC MSI arm-level CNV
tcga_ucec_msi <- tcga[["UCEC.MSI_ArmCNV"]]

# Identify arm columns
arm_cols <- grep("^Arm", colnames(tcga_ucec_msi), value = TRUE)
length(arm_cols)

# Subset arm columns only
tcga_arms_only <- tcga_ucec_msi[, arm_cols, drop = FALSE]

# Compute Global Aneuploidy Score
tcga_ucec_msi_global <- data.frame(
  Patient_ID = rownames(tcga_ucec_msi), 
  Tumor = "UCEC",
  Subtype = "MSI",
  Global_AS = rowSums(abs(tcga_arms_only) > 0.2, na.rm = TRUE)
)

summary(tcga_ucec_msi_global$Global_AS)
```

```{r}
# binarize 
median_AS <- median(tcga_ucec_msi_global$Global_AS, na.rm = TRUE)

tcga_ucec_msi_global$High_Global_AS <- ifelse(
  tcga_ucec_msi_global$Global_AS > median_AS,
  1,
  0
)

table(tcga_ucec_msi_global$High_Global_AS)
prop.table(table(tcga_ucec_msi_global$High_Global_AS))
```

```{r}
# save label file 
write.csv(
  tcga_ucec_msi_global,
  "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/image_analysis/UCEC_MSI_globalAS_labels.csv",
  row.names = FALSE
)
```


```{r}
# Pull Arm1q (continuous) and make Arm_1q_Gain (binary)
arm1q_cont <- tcga_ucec_msi$Arm1q
arm1q_gain <- as.integer(arm1q_cont > 0.2)

df <- tibble(
  Patient_ID = rownames(tcga_ucec_msi),
  Arm1q = arm1q_cont,
  Arm_1q_Gain = arm1q_gain,
  Global_AS = tcga_ucec_msi_global$Global_AS
)


```

Calculate correlation 
```{r}
cor_cont <- cor.test(df$Arm1q, df$Global_AS, method = "spearman", use = "complete.obs")
cor_cont

```

```{r}

# Compare Global_AS between 1q gain vs not (binary vs continuous) ----
df %>%
  group_by(Arm_1q_Gain) %>%
  summarize(
    n = n(),
    median_Global_AS = median(Global_AS, na.rm = TRUE),
    mean_Global_AS = mean(Global_AS, na.rm = TRUE)
  )

wilcox.test(Global_AS ~ Arm_1q_Gain, data = df) 

```

```{r}
# ---- 3) Quick plot (optional) ----
ggplot(df, aes(x = Arm1q, y = Global_AS)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  labs(title = "TCGA UCEC.MSI: Arm1q vs Global Aneuploidy Score",
       x = "Arm1q copy-number value",
       y = "Global_AS (# arms with |CN| > 0.2)")
```


** previous codes using specific arm gain **
# UCEC 1q gain 

```{r}
# TCGA 
tcga_ucec <- tcga[["UCEC_ArmCNV"]]  # Use correct key
names(tcga_ucec)

tcga_labels <- data.frame(
  Patient_ID = paste0(rownames(tcga_ucec), "-tumor"),
  Tumor = "UCEC",
  Arm_1q_Gain = ifelse(tcga_ucec$Arm1q > 0.2, 1, 0)
)

# CPTAC
cptac_ucec <- cptac[cptac$TumorType == "UCEC", ]
cptac_ucec <- rownames_to_column(cptac_ucec, "Patient_ID")

cptac_labels <- cptac_ucec %>%
  mutate(
    Tumor = "UCEC",
    Arm_1q_Gain = ifelse(Arm_1q > 0.2, 1, 0)
  ) %>%
  select(Patient_ID, Tumor, Arm_1q_Gain)

# Combine 
combined_labels <- bind_rows(cptac_labels, tcga_labels)
write.csv(combined_labels, "combined_UCEC_Arm1qGain_labels.csv", row.names = FALSE)
```

```{r}
# Load the two CSV files
df1 <- read.csv("UCEC/arm1q_labels.csv", stringsAsFactors = FALSE)
df2 <- read.csv("UCEC/combined_UCEC_Arm1qGain_labels.csv", stringsAsFactors = FALSE)

# Check the number of rows in each
cat("df1 rows:", nrow(df1), "\n")
cat("df2 rows:", nrow(df2), "\n")

# Check column names
cat("df1 columns:\n"); print(colnames(df1))
cat("df2 columns:\n"); print(colnames(df2))

# Compare Patient_IDs
shared_ids <- intersect(df1$Patient_ID, df2$Patient_ID)
only_in_df1 <- setdiff(df1$Patient_ID, df2$Patient_ID)
only_in_df2 <- setdiff(df2$Patient_ID, df1$Patient_ID)

# Print results
cat("Number of shared Patient_IDs:", length(shared_ids), "\n")
cat("Patient_IDs only in df1 (first 10):\n"); print(head(only_in_df1, 10))
cat("Patient_IDs only in df2 (first 10):\n"); print(head(only_in_df2, 10))
```



```{r}
# Match tiles
tiles <- read.csv("UCEC/tiles.csv")
labels <- read.csv("UCEC/combined_UCEC_Arm1qGain_labels.csv")

# Check overlap
intersect_ids <- intersect(unique(tiles$Patient_ID), labels$Patient_ID)
length(intersect_ids)       # matched
length(unique(tiles$Patient_ID))  # in tiles
length(labels$Patient_ID)         # in label file

# Filter
#tiles_filtered <- tiles %>% filter(Patient_ID %in% labels$Patient_ID)
#write.csv(tiles_filtered, "all_tiles_UCEC_1qgain_filtered.csv", row.names = FALSE)
```

```{r}
# Match tiles
tiles <- read.csv("UCEC/tiles.csv")
labels <- read.csv("UCEC/arm1q_labels.csv")

# Check overlap
intersect_ids <- intersect(unique(tiles$Patient_ID), labels$Patient_ID)
length(intersect_ids)       # matched
length(unique(tiles$Patient_ID))  # in tiles
length(labels$Patient_ID)         # in label file

```

## most aneuploid patient
```{r}
# === TCGA UCEC ===
tcga_ucec <- tcga[["UCEC_ArmCNV"]]
tcga_ucec$Aneuploidy_Score <- apply(tcga_ucec, 1, function(x) sum(abs(x) > 0.2, na.rm = TRUE))
tcga_ucec$Patient_ID <- paste0(rownames(tcga_ucec), "-tumor")

# Split by 1q gain status
tcga_1qgain     <- tcga_ucec[tcga_ucec$Arm1q > 0.2, ]
tcga_not_1qgain <- tcga_ucec[tcga_ucec$Arm1q <= 0.2, ]

# Return both
most_aneuploid_tcga_gain    <- tcga_1qgain[which.max(tcga_1qgain$Aneuploidy_Score), "Patient_ID"]
most_aneuploid_tcga_no_gain <- tcga_not_1qgain[which.max(tcga_not_1qgain$Aneuploidy_Score), "Patient_ID"]

cat("TCGA most aneuploid UCEC with 1q gain: ", most_aneuploid_tcga_gain, "\n")
cat("TCGA most aneuploid UCEC without 1q gain: ", most_aneuploid_tcga_no_gain, "\n")
```

```{r}
# === CPTAC UCEC ===
cptac_ucec <- cptac[cptac$TumorType == "UCEC", ]
cptac_ucec <- tibble::rownames_to_column(cptac_ucec, "Patient_ID")

# Compute aneuploidy score (exclude TumorType and Patient_ID)
cptac_ucec$Aneuploidy_Score <- apply(
  cptac_ucec[, !(colnames(cptac_ucec) %in% c("TumorType", "Patient_ID"))],
  1,
  function(x) sum(abs(x) > 0.2, na.rm = TRUE)
)

# Split by 1q gain
cptac_1qgain     <- cptac_ucec[cptac_ucec$Arm_1q > 0.2, ]
cptac_not_1qgain <- cptac_ucec[cptac_ucec$Arm_1q <= 0.2, ]

# Return both
most_aneuploid_cptac_gain    <- cptac_1qgain[which.max(cptac_1qgain$Aneuploidy_Score), "Patient_ID"]
most_aneuploid_cptac_no_gain <- cptac_not_1qgain[which.max(cptac_not_1qgain$Aneuploidy_Score), "Patient_ID"]

cat("CPTAC most aneuploid UCEC with 1q gain: ", most_aneuploid_cptac_gain, "\n")
cat("CPTAC most aneuploid UCEC without 1q gain: ", most_aneuploid_cptac_no_gain, "\n")
```

