---
title: "R Notebook"
output: html_notebook
---

 

```{r}
# Load packages
library(tidyverse)


# Load data
immune_df <- read.csv("results_immune_allTumors.csv")
survival_df <- read.csv("results_survival.csv")

# Merge on cancer type, cytoband, and CNV status
merged_df <- inner_join(immune_df, survival_df,
                        by = c("OncoTree", "Cytoband", "Comparison"))

# Remove any rows with missing values in key columns
merged_df <- merged_df %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Compute observed Spearman correlation
observed_corr <- cor(merged_df$ImmuneEffect, 
                     merged_df$sign_logP_survival, 
                     method = "spearman")
```

# set up for survival 

```{r}
# Generate survival results using ONCOTREE_CODE to match immune data
cytoband_cols_df <- grep("^[0-9XY]+[pq]", names(df), value = TRUE)
results_survival <- data.frame()

for (tumor in unique(df$ONCOTREE_CODE)) {  # Use ONCOTREE_CODE not CANCER_TYPE
  tumor_data <- df %>% filter(ONCOTREE_CODE == tumor, DRUG_TYPE == "PD-1/PDL-1")
  
  for (band in cytoband_cols_df) {
    for (comparison in c("Gain", "Loss")) {
      subset <- tumor_data %>%
        filter(!!sym(band) %in% c("None", comparison)) %>%
        mutate(CNV = factor(!!sym(band), levels = c("None", comparison)))
      
      if (nrow(subset) >= 10 && length(unique(subset$CNV)) == 2) {
        tryCatch({
          fit <- coxph(Surv(OS_MONTHS, OS) ~ CNV, data = subset)
          hr <- exp(coef(fit))
          p <- summary(fit)$coefficients[5]
          
          results_survival <- rbind(results_survival, data.frame(
            OncoTree = tumor,  # Match naming with immune data
            Cytoband = band,
            Comparison = comparison,
            HR = round(hr, 3),
            P_survival = p,
            log2HR = log2(hr),
            sign_logP_survival = -log10(p) * ifelse(log2(hr) > 0, 1, -1),
            N = nrow(subset)
          ))
        }, error = function(e){})
      }
    }
  }
}


write.csv(results_survival, "results_survival.csv", row.names = FALSE)

```


# Permutation test
Ask if the osberved correlation is stronger than expected by chanc e
- element-wise subtraction 
- merge immune and survival data per cytoband, so each row corresponds to a specific cytoband-tumor comparison combination 
- calcluate correlation across all cytobands using paired values 

```{r}

# Permutation test
set.seed(42)

# shuffles the survival values, breaking the matching between immune & survival 
# recalculates the Spearman correlation 1000x 
# null distribution of correlation values that would happen by random chance 
# p value = count how many times the random correlation was as extreme as your real one 
n_perm <- 1000
perm_corrs <- replicate(n_perm, {
  shuffled <- sample(merged_df$sign_logP_survival)
  cor(merged_df$ImmuneEffect, shuffled, method = "spearman")
})

# Compute p-value (two-sided)
p_value <- mean(abs(perm_corrs) >= abs(observed_corr))

# Print results
cat("Observed correlation:", round(observed_corr, 3), "\n")
cat("Permutation p-value:", p_value, "\n")

# Plot the null distribution
hist(perm_corrs, breaks = 40, col = "gold", border = "black",
     main = paste0("Permutation Test\nObserved corr = ", 
                   round(observed_corr, 3), ", p = ", signif(p_value, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_corr, col = "red", lwd = 2, lty = 2)
```
- very weak but significant inverse correlation between poor survival and immune suprresion (“CNVs associated with immune suppression tend to also show slightly worse survival)
- permuetated 1000 times


Pearson correlation 
```{r}
# Compute observed Pearson correlation
observed_pearson <- cor(merged_df$ImmuneEffect, 
                        merged_df$sign_logP_survival, 
                        method = "pearson")

# Permutation test
set.seed(42)
perm_pearson <- replicate(1000, {
  shuffled <- sample(merged_df$sign_logP_survival)
  cor(merged_df$ImmuneEffect, shuffled, method = "pearson")
})

# Calculate p-value
p_pearson <- mean(abs(perm_pearson) >= abs(observed_pearson))

# Output results
cat("Observed Pearson correlation:", round(observed_pearson, 3), "\n")
cat("Permutation p-value:", p_pearson, "\n")

# Optional: plot null distribution
hist(perm_pearson, breaks = 30, col = "skyblue", main = paste0(
  "Permutation Test (Pearson)\nObserved = ", round(observed_pearson, 3),
  ", p = ", signif(p_pearson, 3)),
  xlab = "Pearson Correlation (permuted)"
)
abline(v = observed_pearson, col = "red", lty = 2, lwd = 2)
```

# cosine similarity 
```{r}
# Cosine similarity = dot(A, B) / (||A|| * ||B||)
cosine_similarity <- function(a, b) {
  sum(a * b) / (sqrt(sum(a^2)) * sqrt(sum(b^2)))
}

# Compute observed cosine similarity
observed_cos <- cosine_similarity(merged_df$ImmuneEffect, merged_df$sign_logP_survival)

# Permutation test (same as before)
perm_cos <- replicate(1000, {
  shuffled <- sample(merged_df$sign_logP_survival)
  cosine_similarity(merged_df$ImmuneEffect, shuffled)
})

# p-value
p_cos <- mean(abs(perm_cos) >= abs(observed_cos))
p_cos
```

# SKCM 
```{r}
# Filter to BLCA only, and remove NAs
skcm_df <- merged_df %>%
  filter(OncoTree == "SKCM") %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Spearman correlation across cytobands
observed_r <- cor(skcm_df$ImmuneEffect, skcm_df$sign_logP_survival, method = "spearman")

# Permutation test
set.seed(42)
perm_r <- replicate(1000, {
  cor(skcm_df$ImmuneEffect, sample(skcm_df$sign_logP_survival), method = "spearman")
})

# P-value
p_val <- mean(abs(perm_r) >= abs(observed_r))

# Output
cat("SKCM — Spearman:", round(observed_r, 3), "p-value:", p_val, "\n")
hist(perm_r, breaks = 30, col = "orange",
     main = paste0("SKCMT — Spearman = ", round(observed_r, 3), ", p = ", signif(p_val, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_r, col = "red", lty = 2, lwd = 2)
```

```{r}
# Filter merged data for SKCM and 1q or 9p cytobands
merged_skcm_1q9p <- merged_df %>%
  filter(OncoTree == "SKCM", grepl("^1q|^9p", Cytoband))

# Remove NAs
merged_skcm_1q9p <- merged_skcm_1q9p %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Rerun Spearman 
observed_spearman <- cor(merged_skcm_1q9p$ImmuneEffect,
                         merged_skcm_1q9p$sign_logP_survival,
                         method = "spearman")

# Step 2: Permutation test
set.seed(42)
perm_spearman <- replicate(1000, {
  shuffled <- sample(merged_skcm_1q9p$sign_logP_survival)
  cor(merged_skcm_1q9p$ImmuneEffect, shuffled, method = "spearman")
})

# Step 3: Compute permutation p-value
p_spearman <- mean(abs(perm_spearman) >= abs(observed_spearman))

# Step 4: Print results
cat("Observed Spearman correlation:", round(observed_spearman, 3), "\n")
cat("Permutation p-value:", p_spearman, "\n")

# Optional: plot null distribution
hist(perm_spearman, breaks = 30, col = "gold",
     main = paste0("Permutation Test (Spearman) — SKCM 1q/9p\nObserved = ", 
                   round(observed_spearman, 3), ", p = ", signif(p_spearman, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_spearman, col = "red", lty = 2, lwd = 2)
```
 - strong inverse relationship: CNVs linked to lower immune infiltration also tend to predict worse survival in melanoma.
 
```{r}
# Filter for SKCM and 1q cytobands
merged_skcm_1q <- merged_df %>%
  filter(OncoTree == "SKCM", grepl("^1q", Cytoband)) %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Spearman correlation
observed_1q <- cor(merged_skcm_1q$ImmuneEffect,
                   merged_skcm_1q$sign_logP_survival,
                   method = "spearman")

# Permutation test
set.seed(42)
perm_1q <- replicate(1000, {
  shuffled <- sample(merged_skcm_1q$sign_logP_survival)
  cor(merged_skcm_1q$ImmuneEffect, shuffled, method = "spearman")
})

# Compute p-value
p_1q <- mean(abs(perm_1q) >= abs(observed_1q))

# Output
cat("SKCM 1q — Spearman:", round(observed_1q, 3), "p-value:", p_1q, "\n")
hist(perm_1q, breaks = 30, col = "gold",
     main = paste0("SKCM 1q — Spearman = ", round(observed_1q, 3), ", p = ", signif(p_1q, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_1q, col = "red", lty = 2, lwd = 2)
```

```{r}
# Filter for SKCM and 9p cytobands
merged_skcm_9p <- merged_df %>%
  filter(OncoTree == "SKCM", grepl("^9p", Cytoband)) %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Spearman correlation
observed_9p <- cor(merged_skcm_9p$ImmuneEffect,
                   merged_skcm_9p$sign_logP_survival,
                   method = "spearman")

# Permutation test
set.seed(42)
perm_9p <- replicate(1000, {
  shuffled <- sample(merged_skcm_9p$sign_logP_survival)
  cor(merged_skcm_9p$ImmuneEffect, shuffled, method = "spearman")
})

# Compute p-value
p_9p <- mean(abs(perm_9p) >= abs(observed_9p))

# Output
cat("SKCM 9p — Spearman:", round(observed_9p, 3), "p-value:", p_9p, "\n")
hist(perm_9p, breaks = 30, col = "gold",
     main = paste0("SKCM 9p — Spearman = ", round(observed_9p, 3), ", p = ", signif(p_9p, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_9p, col = "red", lty = 2, lwd = 2)
```

# Bladder
all cytobands
```{r}
# Filter to BLCA only, and remove NAs
blca_df <- merged_df %>%
  filter(OncoTree == "BLCA") %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Spearman correlation across cytobands
observed_r <- cor(blca_df$ImmuneEffect, blca_df$sign_logP_survival, method = "spearman")

# Permutation test
set.seed(42)
perm_r <- replicate(1000, {
  cor(blca_df$ImmuneEffect, sample(blca_df$sign_logP_survival), method = "spearman")
})

# P-value
p_val <- mean(abs(perm_r) >= abs(observed_r))

# Output
cat("BLCA — Spearman:", round(observed_r, 3), "p-value:", p_val, "\n")
hist(perm_r, breaks = 30, col = "orange",
     main = paste0("BLCA — Spearman = ", round(observed_r, 3), ", p = ", signif(p_val, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_r, col = "red", lty = 2, lwd = 2)
```

```{r}
# Filter for BLCA + 9q only
merged_blca_9q <- merged_df %>%
  filter(OncoTree == "BLCA", grepl("^9q", Cytoband)) %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Spearman correlation
observed_9q <- cor(merged_blca_9q$ImmuneEffect, merged_blca_9q$sign_logP_survival, method = "spearman")

# Permutation test
set.seed(42)
perm_9q <- replicate(1000, {
  shuffled <- sample(merged_blca_9q$sign_logP_survival)
  cor(merged_blca_9q$ImmuneEffect, shuffled, method = "spearman")
})

# p-value
p_9q <- mean(abs(perm_9q) >= abs(observed_9q))

# Output + plot
cat("BLCA 9q — Spearman:", round(observed_9q, 3), "p-value:", p_9q, "\n")
hist(perm_9q, breaks = 30, col = "orange",
     main = paste0("BLCA 9q — Spearman = ", round(observed_9q, 3), ", p = ", signif(p_9q, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_9q, col = "red", lty = 2, lwd = 2)
```

```{r}
# Filter for BLCA + 11p only
merged_blca_11p <- merged_df %>%
  filter(OncoTree == "BLCA", grepl("^11p", Cytoband)) %>%
  filter(!is.na(ImmuneEffect), !is.na(sign_logP_survival))

# Spearman correlation
observed_11p <- cor(merged_blca_11p$ImmuneEffect, merged_blca_11p$sign_logP_survival, method = "spearman")

# Permutation test
set.seed(42)
perm_11p <- replicate(1000, {
  shuffled <- sample(merged_blca_11p$sign_logP_survival)
  cor(merged_blca_11p$ImmuneEffect, shuffled, method = "spearman")
})

# p-value
p_11p <- mean(abs(perm_11p) >= abs(observed_11p))

# Output + plot
cat("BLCA 11p — Spearman:", round(observed_11p, 3), "p-value:", p_11p, "\n")
hist(perm_11p, breaks = 30, col = "orange",
     main = paste0("BLCA 11p — Spearman = ", round(observed_11p, 3), ", p = ", signif(p_11p, 3)),
     xlab = "Spearman Correlation (permuted)")
abline(v = observed_11p, col = "red", lty = 2, lwd = 2)
```
 
# Correlation plot 
```{r}
# Merge on OncoTree, Cytoband, Comparison
merged_df <- inner_join(immune_df, survival_df,
                        by = c("OncoTree", "Cytoband", "Comparison"))

# Compute flipped signed -log10(p) for immune and survival
# NOTE: Immune sign is flipped to align with survival (so negative immune becomes positive)
merged_df <- merged_df %>%
  mutate(
    sign_logP_immune = -1 * (-log10(P_immune) * sign(ImmuneEffect)), # flip sign 
    sign_logP_surv   = -log10(P_survival) * sign(log2HR)
  ) %>%
  filter(!is.na(sign_logP_immune), !is.na(sign_logP_surv))

# Format p-value display
format_p <- function(p) {
  ifelse(p < 0.001, "< 0.001", signif(p, 2))
}

# Compute Spearman correlation
cor_val <- cor(merged_df$sign_logP_immune, merged_df$sign_logP_surv, method = "spearman")
p_val <- cor.test(merged_df$sign_logP_immune, merged_df$sign_logP_surv, method = "spearman")$p.value

# Plot correlation
ggplot(merged_df, aes(x = sign_logP_immune, y = sign_logP_surv)) +
  geom_point(alpha = 0.7, color = "#4F6DB8") +
  geom_smooth(method = "lm", color = "darkred", se = FALSE, linetype = "dashed") +
  labs(
    title = "Immune vs Survival Association Across Cytobands",
    subtitle = paste0("Spearman rho = ", round(cor_val, 3), ", p ", format_p(p_val)),
    x = "Immune Score (signed -log10(p))",
    y = "Survival  (signed -log10(p))"
  ) +
  theme_minimal(base_size = 13)
```

```{r}
# Tumor-specific correlation plots
tumors <- c("SKCM", "LUAD", "BLCA", "COAD", "HNSC")

for (tumor in tumors) {
  df_tumor <- merged_df %>% filter(OncoTree == tumor)
  
  cor_val <- cor(df_tumor$sign_logP_immune, df_tumor$sign_logP_surv, method = "spearman")
  p_val <- cor.test(df_tumor$sign_logP_immune, df_tumor$sign_logP_surv, method = "spearman")$p.value
  
  p <- ggplot(df_tumor, aes(x = sign_logP_immune, y = sign_logP_surv)) +
    geom_point(alpha = 0.7, color = "#4F6DB8") +
    geom_smooth(method = "lm", se = FALSE, color = "darkred", linetype = "dashed") +
    labs(
      title = paste(tumor, "Immune vs Survival Association"),
      subtitle = paste0("Spearman ρ = ", round(cor_val, 3),
                        ", p = ", format_p(p_val)),
      x = "Immune Score (signed -log10(p))",
      y = "Survival (signed -log10(p))"
    ) +
    theme_minimal(base_size = 13)
  
  print(p)
}
```

```{r}
# Tumor-specific correlation plots with cytoband labels on top right
library(ggrepel)

tumors <- c("SKCM", "LUAD", "BLCA")

for (tumor in tumors) {
  df_tumor <- merged_df %>% filter(OncoTree == tumor)
  
  # Calculate Spearman correlation
  cor_val <- cor(df_tumor$sign_logP_immune, df_tumor$sign_logP_surv, method = "spearman")
  p_val <- cor.test(df_tumor$sign_logP_immune, df_tumor$sign_logP_surv, method = "spearman")$p.value
  
  # Get top 5 cytobands near top-right (strongest joint positive associations)
  df_tumor <- df_tumor %>%
    mutate(distance = sqrt(sign_logP_immune^2 + sign_logP_surv^2)) %>%
    arrange(desc(sign_logP_immune + sign_logP_surv))  # prioritize top right corner
  
  top_cytobands <- df_tumor %>%
    slice_max(order_by = sign_logP_immune + sign_logP_surv, n = 10)
  
  # Plot
  p <- ggplot(df_tumor, aes(x = sign_logP_immune, y = sign_logP_surv)) +
    geom_point(alpha = 0.7, color = "#4F6DB8") +
    geom_text_repel(data = top_cytobands, aes(label = Cytoband),
                    size = 3.5, color = "black", max.overlaps = Inf) +
    geom_smooth(method = "lm", se = FALSE, color = "darkred", linetype = "dashed") +
    labs(
      title = paste(tumor, "Immune vs Survival Association"),
      subtitle = paste0("Spearman ρ = ", round(cor_val, 3),
                        ", p = ", ifelse(p_val < 0.001, "< 0.001", signif(p_val, 2))),
      x = "Immune Score (signed -log10(p))",
      y = "Survival (signed -log10(p))"
    ) +
    theme_minimal(base_size = 13)
  
  print(p)
}
```
For each tumor type, separate gain and losses 
make color proportional to strength of correlation (gradient) --> grey to blue to for losses and grey to red for gains 
```{r}
tumors <- c("SKCM", "BLCA", "LUAD")

for (tumor in tumors) {
  df_tumor <- merged_df %>%
    filter(OncoTree == tumor,
           is.finite(ImmuneEffect),
           is.finite(sign_logP_survival))

  for (comparison in c("Gain", "Loss")) {
    df_sub <- df_tumor %>%
      filter(Comparison == comparison) %>%
      mutate(signal_strength = abs(ImmuneEffect + sign_logP_survival))

    if (nrow(df_sub) < 5) next

    cor_val <- cor(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")
    p_val  <- cor.test(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")$p.value

    color_gradient <- if (comparison == "Gain") {
      scale_color_gradient(low = "grey80", high = "#EA6A6A")
    } else {
      scale_color_gradient(low = "grey80", high = "#4F6DB8")
    }

    p <- ggplot(df_sub, aes(x = ImmuneEffect, y = sign_logP_survival, color = signal_strength)) +
      geom_point(size = 3, alpha = 0.85) +
      geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
      color_gradient +
      labs(
        title = paste(tumor, "-", comparison),
        subtitle = paste0("Spearman ρ = ", round(cor_val, 3),
                          ", p = ", ifelse(p_val < 0.001, "< 0.001", signif(p_val, 2))),
        x = "Immune Score (signed -log10(p))",
        y = "Survival Score (signed -log10(p))",
        color = "Signal Strength"
      ) +
      theme_minimal(base_size = 13)

    print(p)
  }
}
```

```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)

tumors <- c("SKCM", "BLCA", "LUAD")

for (tumor in tumors) {
  
  df_tumor <- merged_df %>%
    filter(OncoTree == tumor,
           is.finite(ImmuneEffect),
           is.finite(sign_logP_survival))
  
  for (comparison in c("Gain", "Loss")) {
    
    df_sub <- df_tumor %>%
      filter(Comparison == comparison) %>%
      mutate(signal_strength = abs(ImmuneEffect) + abs(sign_logP_survival))
    
    if (nrow(df_sub) < 5) next
    
    cor_val <- cor(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")
    p_val  <- cor.test(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")$p.value
    
    top_labels <- df_sub %>%
      slice_max(order_by = signal_strength, n = 10)
    
    color_gradient <- if (comparison == "Gain") {
      scale_color_gradient(low = "grey80", high = "#EA6A6A")
    } else {
      scale_color_gradient(low = "grey80", high = "#4F6DB8")
    }
    
    p <- ggplot(df_sub, aes(x = ImmuneEffect,
                            y = sign_logP_survival,
                            color = signal_strength)) +
      geom_point(size = 3, alpha = 0.85) +
      geom_text_repel(data = top_labels, aes(label = Cytoband),
                      color = "black", size = 3.5, max.overlaps = Inf) +
      geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
      color_gradient +
      labs(
        title = paste(tumor, "-", comparison),
        subtitle = paste0("Spearman ρ = ", round(cor_val, 3),
                          ", p = ", ifelse(p_val < 0.001, "< 0.001", signif(p_val, 2))),
        x = "Immune Score",
        y = "Survival Score (signed -log10(p))",
        color = "Signal Strength"
      ) +
      theme_minimal(base_size = 13)
    
    print(p)
  }
}
```
```{r}
# save all plots to 1 pdf
# Save all plots to one PDF file
pdf("immune_survival_correlation_plots.pdf", width = 8, height = 6)

# Your existing plotting code starts here
library(ggplot2)
library(ggrepel)
library(dplyr)

tumors <- c("SKCM", "BLCA", "LUAD")

for (tumor in tumors) {
  
  df_tumor <- merged_df %>%
    filter(OncoTree == tumor,
           is.finite(ImmuneEffect),
           is.finite(sign_logP_survival))
  
  for (comparison in c("Gain", "Loss")) {
    
    df_sub <- df_tumor %>%
      filter(Comparison == comparison) %>%
      mutate(signal_strength = abs(ImmuneEffect) + abs(sign_logP_survival))
    
    if (nrow(df_sub) < 5) next
    
    cor_val <- cor(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")
    p_val  <- cor.test(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")$p.value
    
    top_labels <- df_sub %>%
      slice_max(order_by = signal_strength, n = 10)
    
    color_gradient <- if (comparison == "Gain") {
      scale_color_gradient(low = "grey80", high = "#EA6A6A")
    } else {
      scale_color_gradient(low = "grey80", high = "#4F6DB8")
    }
    
    p <- ggplot(df_sub, aes(x = ImmuneEffect,
                            y = sign_logP_survival,
                            color = signal_strength)) +
      geom_point(size = 3, alpha = 0.85) +
      geom_text_repel(data = top_labels, aes(label = Cytoband),
                      color = "black", size = 3.5, max.overlaps = Inf) +
      geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
      color_gradient +
      labs(
        title = paste(tumor, "-", comparison),
        subtitle = paste0("Spearman ρ = ", round(cor_val, 3),
                          ", p = ", ifelse(p_val < 0.001, "< 0.001", signif(p_val, 2))),
        x = "Immune Score",
        y = "Survival Score (signed -log10(p))",
        color = "Signal Strength"
      ) +
      theme_minimal(base_size = 13)
    
    print(p)  # This sends the plot to the PDF
  }
}

# Close the PDF file
dev.off()
```

```{r}
# save all plots to 1 pdf
# Save all plots to one PDF file
pdf("immune_survival_COAD_HNSC.pdf", width = 8, height = 6)

# Your existing plotting code starts here
library(ggplot2)
library(ggrepel)
library(dplyr)

tumors <- c("COAD", "HNSC")

for (tumor in tumors) {
  
  df_tumor <- merged_df %>%
    filter(OncoTree == tumor,
           is.finite(ImmuneEffect),
           is.finite(sign_logP_survival))
  
  for (comparison in c("Gain", "Loss")) {
    
    df_sub <- df_tumor %>%
      filter(Comparison == comparison) %>%
      mutate(signal_strength = abs(ImmuneEffect) + abs(sign_logP_survival))
    
    if (nrow(df_sub) < 5) next
    
    cor_val <- cor(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")
    p_val  <- cor.test(df_sub$ImmuneEffect, df_sub$sign_logP_survival, method = "spearman")$p.value
    
    top_labels <- df_sub %>%
      slice_max(order_by = signal_strength, n = 10)
    
    color_gradient <- if (comparison == "Gain") {
      scale_color_gradient(low = "grey80", high = "#EA6A6A")
    } else {
      scale_color_gradient(low = "grey80", high = "#4F6DB8")
    }
    
    p <- ggplot(df_sub, aes(x = ImmuneEffect,
                            y = sign_logP_survival,
                            color = signal_strength)) +
      geom_point(size = 3, alpha = 0.85) +
      geom_text_repel(data = top_labels, aes(label = Cytoband),
                      color = "black", size = 3.5, max.overlaps = Inf) +
      geom_smooth(method = "lm", se = FALSE, linetype = "dashed", color = "black") +
      color_gradient +
      labs(
        title = paste(tumor, "-", comparison),
        subtitle = paste0("Spearman ρ = ", round(cor_val, 3),
                          ", p = ", ifelse(p_val < 0.001, "< 0.001", signif(p_val, 2))),
        x = "Immune Score",
        y = "Survival Score (signed -log10(p))",
        color = "Signal Strength"
      ) +
      theme_minimal(base_size = 13)
    
    print(p)  # This sends the plot to the PDF
  }
}

# Close the PDF file
dev.off()
```

