---
title: "timing_mutation_gain"
author: "Annabel Chen"
date: "2026-02-16"
output: html_document
---

Gain timing 
- read in a gain timing index table 
- for each CNV segment/gene, compare mutation multiplicity vs major copy number
- output timing score (index) per gene per cancer type 
  - constructed as how often the SCNA happened before mutations 
  - If across loci in that region, SCNA is inferred to precede the mutations a lot of time, the region get a high timing index = early event
  - if it rarely precedes mutations, it is later 

```{r}
# library + paths
library(tidyverse)
library(AnnotationDbi)
library(org.Hs.eg.db)

gao_dir     <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/timing_cancer_events/EachGene"
summary_dir <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab"
reg_file    <- "/Users/annabelshinichen/Desktop/SchoolWork/MSBMI/Davoli_lab/timing_cancer_events/Regression_Outputs_is0.3_ArmCNV_continuous0.2.rds"
```


```{r}
# reference: list all Gao RData files
gao_files <- list.files(gao_dir, pattern = "\\.Rdata$", full.names = TRUE)
length(gao_files)
head(gao_files, 10)
```

Load timing file 
	•	rows = ENSG genes
	•	columns = TCGA samples
	•	values = timing index (high = early, low = late)
	
Compute a single score per gene per tumor type 
- use mean or median 
- collapse to 1 timing score per gene per cancer type 


```{r}
# Load Gao timing
gao_gene_timing <- list()

for (file_path in gao_files) {
  tumor <- tools::file_path_sans_ext(basename(file_path))  # e.g. "LUAD"

  e <- new.env(parent = emptyenv())
  load(file_path, envir = e)
  if (!exists("time_index_matrix", envir = e)) next

  mat <- e$time_index_matrix
  timing_vec <- rowMeans(mat, na.rm = TRUE)

  gao_gene_timing[[tumor]] <- tibble(
    ENSG = names(timing_vec),
    TimingScore = as.numeric(timing_vec)
  )
}

stopifnot(length(gao_gene_timing) > 0)
names(gao_gene_timing)[1:5]

```

Build a SYMBOL → ENSG mapping (for ALL genes in your summary tables)
```{r}
setwd(summary_dir)
data_table <- readRDS("SummaryTables/TCGA_summary_tables.RDS")

all_symbols <- sort(unique(unlist(lapply(data_table, function(df) df$Gene))))

raw_map <- AnnotationDbi::select(
  x       = org.Hs.eg.db,
  keys    = all_symbols,
  keytype = "SYMBOL",
  columns = c("SYMBOL", "ENSEMBL")
)

raw_map <- raw_map[!is.na(raw_map$ENSEMBL), c("SYMBOL","ENSEMBL")]
raw_map <- raw_map[!duplicated(raw_map$SYMBOL), ]  # keep ONE ENSG per symbol

symbol_to_ensg <- tibble(
  Gene = raw_map$SYMBOL,
  ENSG = raw_map$ENSEMBL
)
```

# Arm

```{r}
# arm timing per tumor
arm_timing <- list()

common_tumors <- intersect(names(data_table), names(gao_gene_timing))

for (tumor in common_tumors) {

  # summary table: Gene SYMBOL -> Cytoband
  tmp <- data_table[[tumor]] %>%
    left_join(symbol_to_ensg, by = "Gene")

  # add Gao timing by ENSG
  tmp <- tmp %>%
    left_join(gao_gene_timing[[tumor]], by = "ENSG")

  # arm from cytoband
  tmp <- tmp %>%
    mutate(Arm = sub("([0-9XY]+[pq]).*", "\\1", Cytoband))

  # arm-level mean timing
  arm_timing[[tumor]] <- tmp %>%
    group_by(Arm) %>%
    summarize(
      GaoArmTiming = mean(TimingScore, na.rm = TRUE),
      n_genes = sum(!is.na(TimingScore)),
      .groups = "drop"
    )
}

# make a long timing table
arm_reg <- readRDS(reg_file)
stopifnot(is.list(arm_reg), length(arm_reg) > 0)
names(arm_reg)[1:10]

arm_reg_df <- purrr::imap_dfr(arm_reg, function(sub, tt) {
  sub <- as.data.frame(sub)

  need <- c("Arm", "Event", "Z_val", "P_val")
  if (!all(need %in% colnames(sub))) return(tibble::tibble())

  sub %>%
    dplyr::filter(Event == "Gain") %>%
    dplyr::mutate(TCGA_TumorType = tt) %>%
    dplyr::select(TCGA_TumorType, Arm, Z_val, P_val)
})

# make long table: TumorType + Arm + GaoArmTiming
gao_arm_df <- purrr::imap_dfr(arm_timing, ~dplyr::mutate(.x, TCGA_TumorType = .y)) %>%
  dplyr::select(TCGA_TumorType, Arm, GaoArmTiming, n_genes)

head(gao_arm_df)
```

 Load regression output and make GAINS-ONLY long table (creates arm_reg_df)
```{r}
# Join Gao arm timing onto regression results ----
merged_gain <- arm_reg_df %>%
  dplyr::left_join(gao_arm_df, by = c("TCGA_TumorType","Arm")) %>%
  dplyr::mutate(
    signed_logp = sign(Z_val) * -log10(P_val)
  )

# sanity checks
c(
  n_rows = nrow(merged_gain),
  n_with_gao = sum(!is.na(merged_gain$GaoArmTiming)),
  frac_with_gao = mean(!is.na(merged_gain$GaoArmTiming))
)

```
 
```{r}
cor_df <- merged_gain %>%
  dplyr::group_by(TCGA_TumorType) %>%
  dplyr::summarize(
    n_complete = sum(!is.na(GaoArmTiming) & !is.na(signed_logp)),
    spearman_r = ifelse(
      n_complete >= 3,
      cor(GaoArmTiming, signed_logp, method = "spearman", use = "complete.obs"),
      NA_real_
    ),
    .groups = "drop"
  ) %>%
  dplyr::arrange(desc(spearman_r))

cor_df
```
 
```{r}
cor_df %>% dplyr::filter(is.na(spearman_r)) %>% dplyr::arrange(n_complete)
```
 
## plot
```{r}
library(ggpubr)
library(ggrepel)

plot_df <- merged_gain %>%
  dplyr::filter(!is.na(GaoArmTiming), !is.na(signed_logp)) %>%
  dplyr::mutate(
    ChrArm = factor(ifelse(grepl("p$", Arm), "p", "q"), levels = c("p","q")),
    Chromosome = factor(gsub("p|q", "", Arm), levels = c(as.character(1:22), "X"))
  )

p_scatter <- ggscatter(
  plot_df,
  x = "GaoArmTiming",
  y = "signed_logp",
  shape = "ChrArm",
  color = "Chromosome",
  xlab = "Gao gain timing (higher = earlier)",
  ylab = "Signed –log10(P) (immune association)",
  size = 2,
  alpha = 0.9
) +
  facet_wrap(~ TCGA_TumorType) +
  geom_hline(yintercept = 0, alpha = 0.25) +
  geom_smooth(method = "lm", color = "black", linewidth = 0.4, se = FALSE) +
  stat_cor(method = "spearman", size = 3, na.rm = TRUE) +
  ggrepel::geom_text_repel(
    aes(label = Arm),
    size = 2.2,
    max.overlaps = 30
  ) +
  theme_bw()

p_scatter
```
 
```{r}
# Plot scatter plot for only BRCA
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggrepel)

# 1) Gao arm timing for BRCA + BLCA
gao_bb <- gao_arm_df %>%
  dplyr::filter(TCGA_TumorType %in% c("BRCA", "BLCA")) %>%
  dplyr::select(TCGA_TumorType, Arm, GaoArmTiming)

# 2) Regression rows for BRCA + BLCA (overall only; excludes BRCA.neg/pos etc)
reg_bb <- arm_reg_df %>%
  dplyr::filter(TCGA_TumorType %in% c("BRCA", "BLCA")) %>%
  dplyr::mutate(
    signed_logp = sign(Z_val) * -log10(P_val)
  )

# 3) Join within tumor type + arm
plot_df <- reg_bb %>%
  dplyr::left_join(gao_bb, by = c("TCGA_TumorType", "Arm")) %>%
  dplyr::filter(!is.na(GaoArmTiming), !is.na(signed_logp)) %>%
  dplyr::mutate(
    ChrArm = factor(ifelse(grepl("p$", Arm), "p", "q"), levels = c("p","q")),
    label = Arm
  )

# 4) Plot: two panels (BRCA and BLCA)
p <- ggplot(plot_df, aes(x = GaoArmTiming, y = signed_logp)) +
  geom_point(aes(shape = ChrArm), size = 2, alpha = 0.9, color = "black") +
  geom_hline(yintercept = 0, alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.4) +
  ggpubr::stat_cor(method = "spearman", label.x.npc = 0.02, label.y.npc = 0.95, size = 3, na.rm = TRUE) +
  ggrepel::geom_text_repel(aes(label = label), size = 2.6, max.overlaps = 40) +
  facet_wrap(~TCGA_TumorType, nrow = 1) +
  theme_bw(base_size = 12) +
  labs(
    x = "Timing",
    y = "Signed -log10(P)"
  )

print(p)
```
 
```{r}
# Plot bar plot
p_bar <- cor_df %>%
  dplyr::filter(!is.na(spearman_r)) %>%
  dplyr::mutate(TCGA_TumorType = forcats::fct_reorder(TCGA_TumorType, spearman_r)) %>%
  ggplot(aes(x = TCGA_TumorType, y = spearman_r)) +
  geom_col() +
  coord_flip() +
  theme_bw(base_size = 13) +
  labs(
    title = "Correlation (Gain iiming vs Immune Invasion)",
    x = "Tumor type",
    y = "Spearman r"
  )

p_bar
```

 